---
title: "comparing_results"
author: "Lukas"
date: "2025-12-04"
output: html_document
params:
  config: null
---

```{r setup, include=FALSE}

#===============================================================================
# Run setup 
#===============================================================================

print(.libPaths()) # Debugging; can be commented out.

#-------------------------------------------------------------------------------
# Find a specific file by walking up parent directories
#-------------------------------------------------------------------------------

find_up_file <- function(start_dir, filename, max_up = 15) {
  d <- normalizePath(start_dir, mustWork = TRUE)

  for (i in 0:max_up) {
    candidate <- file.path(d, filename)
    if (file.exists(candidate)) return(candidate)

    parent <- dirname(d)
    if (identical(parent, d)) break
    d <- parent
  }

  stop("Could not find '", filename, "' searching upward from: ", start_dir)
}

#-------------------------------------------------------------------------------
# Determine project root
#-------------------------------------------------------------------------------

# Now we find the location of the Rmd file we are running.
# This can be set externally via SOURCE_RMD.
rmd_path <- Sys.getenv("SOURCE_RMD", unset = "")

start_dir <- if (nzchar(rmd_path) && file.exists(rmd_path)) {
  dirname(normalizePath(rmd_path, mustWork = TRUE))
} else {
  getwd()
}

# Anchor file used to identify project root.
# compare_results.Rmd should live next to, or below, data_analysis_with_MAUDE.Rmd.
anchor_rmd <- find_up_file(start_dir, "data_analysis_with_MAUDE.Rmd", max_up = 1)

# The anchor's directory is the project root that contains functions/, config.yaml, etc.
project_root_dir <- dirname(anchor_rmd)


#-------------------------------------------------------------------------------
# Source pipeline functions
#-------------------------------------------------------------------------------

source(file.path(project_root_dir, "R/R_functions", "s00_bootstrap.R"))
bootstrap_pipeline(project_root_dir)

#-------------------------------------------------------------------------------
# Run setup
#-------------------------------------------------------------------------------

cfg <- project_setup(
  project_root_dir,
  config_path = file.path(project_root_dir, "config.yaml")
)
cfg <- readRDS("/g/steinmetz/link/Amplicon_barcode_analysis/Jia_CM_prelim/.pipeline_state/resolved_config.rds")


################################################################################
############# LOAD COMPARISON SPECIFIC LIBS #####################################
################################################################################

library(VennDiagram)
library(grid)
library(gridExtra)
library(clusterProfiler)
library(ggbreak)
library(STRINGdb)
library(igraph)
library(GO.db)
library(scales)
```

``` {r load_comparision_specific_libs, eval=TRUE}
# library(VennDiagram)
# library(grid)
# library(gridExtra)
# library(clusterProfiler)
# library(ggbreak)
# library(STRINGdb)
# library(igraph)
# library(GO.db)
# library(scales)
```
# The following codeblock exists to do some basic familiarization with MAUDE,
and to troubleshoot it if things fail. You can ignore it.  
```{r make_example_df, eval=FALSE}
if (FALSE){
  getAnywhere(findGuideHitsAllScreens)
  getAnywhere(getElementwiseStats)
  getAnywhere(findGuideHits)
  getAnywhere(getZScalesWithNTGuides())
  getAnywhere(optim())
}


# Create the data for maude
maude_example_df_1 <- data.frame(sgRNA = c(LETTERS[1:6], "CONTROL_C_NONTARG_G", "CONTROL_C_NONTARG_H", "CONTROL_C_I", "CONTROL_C_J"),
                               input = c(4.1, 3, 5, 3, 8, 7, 5, 6, 7, 5),
                               upper = c(1, 9, 5, 0, 4, 2, 0, 1, 3, 2),
                               lower = c(4, 4, 1, 1, 3, 1, 0, 2, 2, 1),
                               sublib = "sublib_1")
maude_example_df_2 <- maude_example_df_1 %>%
  mutate(input = input + 1,
         upper = upper + 1,
         lower = lower + 1,
         exp = 'rep1',
         isNontargeting = case_when(str_detect(sgRNA, "^CONTROL_C_NONTARG_") ~ FALSE, # Anything else → FALSE
                                    str_detect(sgRNA, "^CONTROL_C_") ~ TRUE, # CONTROL_C_X → TRUE
                                    TRUE ~ FALSE # Anything else → FALSE) %>%
                                    )
  )

# Define bin stats with 10% for lower/upper each
lower_bin_end = 0.1
upper_bin_start = 1 - 0.1

maude_example_bins <- tibble(
  Bin = rep(c("upper", "lower"), length(unique(maude_example_df_2[["exp"]]))),  # Repeat 'upper' and 'lower' for each exp
  exp = rep(unique(maude_example_df_2[["exp"]]), each = 2),  # Repeat each exp value twice
  binStartQ = ifelse(rep(c("upper", "lower"), length(unique(maude_example_df_2[["exp"]]))) == "lower", 0.001, upper_bin_start),
  binEndQ = ifelse(rep(c("upper", "lower"), length(unique(maude_example_df_2[["exp"]]))) == "lower", lower_bin_end, 0.999),
  fraction = binEndQ - binStartQ,
  binStartZ = qnorm(binStartQ),
  binEndZ = qnorm(binEndQ)
) %>%
  select(Bin, binStartQ, binEndQ, fraction, binStartZ, binEndZ, exp) %>%
  as.data.frame()


maude_example_guide_stats <- findGuideHitsAllScreens(
  experiments = unique(maude_example_df_2['exp']),
  countDataFrame = maude_example_df_2,
  binStats = maude_example_bins,
  sortBins = c('lower', 'upper'),
  unsortedBin = 'input',
  negativeControl = 'isNontargeting'
)
```

## Generate the long format read/umi counts either from my pipline or from bcwithqcs
Also provides information for each file about coverage and correct alignment

```{r get_count_df_long, eval=TRUE}
process_bcwithqc_data_backup <- process_bcwithqc_data
pseudocount_added <- FALSE

sink(get_file_path(rds_output_folder,paste0(file_info_suffix,"_coverage.txt")),
     split = TRUE) # capture output into a file

if (read_counting == "bcwithqc"){
  count_df_long <- process_bcwithqc_data(bcwithqc_dir,
                                     data_type = data_type,
                                     merged_sgRNA_df = merged_sgRNA_df,
                                     skip_list = skip_list)
  if (extra_suffix == "rf"){
      count_df_long <- count_df_long %>% 
      mutate(count = ifelse(count < simplified_rf_threshold, 0, count))        
      }
} else {
  if (read_counting == "align_UMI_tools"){
    if (data_type == "umis"){
      count_df_long <- process_folder_files(dedup_output_folder,
                                            skip_list = skip_list) #Add threshold df if thresholds should be applied 
    }
    if (data_type == "reads"){
      count_df_long <- process_folder_files(mapped_output_folder,
                                            skip_list = skip_list) #Add threshold df if thresholds should be applied 
      if (extra_suffix == "rf"){
        count_df_long <- count_df_long %>% 
        mutate(count = ifelse(count < simplified_rf_threshold, 0, count))        
      }

    }
       
  } else {
    if (read_counting == "bcwithqc_rf"){
      count_df_long <- process_bcwithqc_rf_data(bcwithqc_rf_folder,
                                            data_type = data_type,
                                            merged_sgRNA_df = merged_sgRNA_df,
                                            skip_list = skip_list)
      
      process_bcwithqc_data <- process_bcwithqc_rf_data
      if (extra_suffix == "rf"){
        count_df_long <- count_df_long %>% 
        mutate(count = ifelse(count < simplified_rf_threshold, 0, count))        
      }
    } else {
      stop("Unsupported read_counting specified. Exiting. Pipline must be bcwithqc, bcwithqc_rf or align_UMI_tools.")
    }
  }  
}
sink()
if (extra_suffix == "cf"){
  count_df_long <- count_df_long %>%
    filter(!(group != "targeting" & count < simplified_cf_threshold))
}

# Extend the include_controls_list if use_only_these_controls_list is given
if (exists("use_only_these_controls_list")) {
  if (length(use_only_these_controls_list) > 0){
    # list of sgRNAs that were not targeting (before) and not in allowed controls
    excluded_controls <- count_df_long %>%
      filter(group_category != "targeting", !sgRNA %in% use_only_these_controls_list) %>%
      distinct(sgRNA) %>%
      pull(sgRNA)
  
    include_controls_list <- c(include_controls_list, excluded_controls)
  }
}
if (length(include_controls_list) > 0){
  for (control_gene in include_controls_list){
    count_df_long$group_category[grepl(control_gene, count_df_long$sgRNA)] <- "targeting"
  }
}

```

```{r prepare_data_for_maude, eval=TRUE}
if (subsample_controls == TRUE){
  # count_df_long_old <- subsample_controls_func_old(count_df_long, merged_sgRNA_df)
  count_df_long <- subsample_controls_func(count_df_long, merged_sgRNA_df)
}

if (!(norm_method %in% c("","control_median"))){
  stop("Error: 'norm_method' must be one of '', 'control_median'. The script will now stop.")
}
if (norm_method == "control_median"){
  count_df_long <- normalize_count_df_long(count_df_long,
                                           norm_method = norm_method)
}


  
maude_counts_df <- count_df_long_to_wide(count_df_long = count_df_long,
                                         print = FALSE,
                                         drop_0s = drop_0s,
                                         recover_input = recover_input)
if (!(method %in% c("","rep","sum",'rep_sample','rep_sublib'))){
  stop("Error: 'method' must be one of '', 'rep', 'rep_sample', 'rep_sublib', or 'sum'. The script will now stop.")
}

if (method == ""){
  maude_counts_df <- maude_counts_df %>% 
    mutate(exp = "rep1")
}
if (method == "rep"){

}
if (method == "rep_sample"){
  maude_counts_df <- maude_counts_df %>%
    mutate(exp = sample)
}
if (method == "rep_sublib"){
  maude_counts_df <- maude_counts_df %>%
    mutate(exp = sublib)
}
if (method == "sum"){
  stop("Method 'sum' is deprecated, use the option combine_for_guide_stats instead")
  maude_counts_df <- count_df_long_to_wide(count_df_long = count_df_long,
                                         print = FALSE,
                                         drop_0s = drop_0s,
                                         recover_input = TRUE,
                                         for_sum = TRUE)
  # Group by sgRNA and summarize the required columns
  maude_counts_df <- maude_counts_df %>%
    group_by(sgRNA, sublib) %>%
    summarize(
      input = pmax(sum(input, na.rm = TRUE), 0),  # Sum and ensure minimum is 0
      upper = pmax(sum(upper, na.rm = TRUE), 0),  # Sum and ensure minimum is 0
      lower = pmax(sum(lower, na.rm = TRUE), 0),  # Sum and ensure minimum is 0
      isNontargeting = dplyr::first(isNontargeting),  # Take the first value of isNontargeting (same for all in the group)
      .groups = 'drop'  # Drop the group structure after summarizing
    ) %>%
    mutate(
      exp = "rep1",
      input = input + 1,
      upper = upper + 1,
      lower = lower + 1
    )
}

if (strict_mode){
  if (pseudocount_added){
    umi_threshold <- 2
  } else {
    umi_threshold <- 1
  } 
  cat("Strict Mode enabled\n")
  cat("Rows before strict mode: \t", nrow(maude_counts_df),"\n")
  maude_counts_df <- maude_counts_df %>%
    filter(if_any(c(input, upper, lower), ~ . <= umi_threshold))
  cat("Rows after strict mode: \t", nrow(maude_counts_df),"\n")
}
if (length(include_controls_list) > 0) {
  for (control_gene in include_controls_list) {
    maude_counts_df$isNontargeting[grepl(control_gene, maude_counts_df$sgRNA)] <- FALSE
  }
}
if (exists("use_only_these_controls_list")) {
  if (length(use_only_these_controls_list) > 0){
    maude_counts_df$isNontargeting[ !(maude_counts_df$sgRNA %in% use_only_these_controls_list) ] <- FALSE
  }
}
if (combine_for_guide_stats != ""){
  if (combine_for_guide_stats == "sample"){
    maude_counts_df <- maude_counts_df %>%
      group_by(sgRNA, sublib) %>%
      summarise(
        # set sample name for the combined rows
        sample = "sample_1",
        
        # add up integer count columns
        input = sum(input, na.rm = TRUE),
        lower = sum(lower, na.rm = TRUE),
        upper = sum(upper, na.rm = TRUE),
        
        # keep the first value for everything else
        exp = dplyr::first(exp),
        isNontargeting = dplyr::first(isNontargeting),
        
        # keep one sublib/sgRNA (also fine even though they're grouping keys)
        .groups = "drop"
      )
  }
  if (combine_for_guide_stats == "sublib"){
    maude_counts_df <- maude_counts_df %>%
      group_by(sgRNA, sample) %>%
      summarise(
        # set sample name for the combined rows
        sublib = "sublib_1",
        
        # add up integer count columns
        input = sum(input, na.rm = TRUE),
        lower = sum(lower, na.rm = TRUE),
        upper = sum(upper, na.rm = TRUE),
        
        # keep the first value for everything else
        exp = dplyr::first(exp),
        isNontargeting = dplyr::first(isNontargeting),
        
        # keep one sublib/sgRNA (also fine even though they're grouping keys)
        .groups = "drop"
      )    
  }
}  
```
## Some exploration options for the raw reads/umi counts
Mostly for comparing this read_counting to bcwithqcs bcwithqc, ignore this. 
```{r basic_read_counting_comparision, eval=FALSE}


PA_count_diff_df <- compare_bcwithqc_align_UMI_tools(
  data_type = "reads",
  cfg = cfg,
  bcwithqc_dir = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_bcwithqc/bcwithqc_output",
  align_UMI_tools_read_folder = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_align/mapped",
  read_diff_histo_x_lim = c(-1000,10),
  read_diff_histo_bin_width = 10
  )
DCA_count_diff_df <- compare_bcwithqc_align_UMI_tools(
  data_type = "reads",
  cfg = cfg,
  bcwithqc_dir = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_bcwithqc/bcwithqc_output",
  align_UMI_tools_read_folder = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_align/mapped",
  read_diff_histo_x_lim = c(-1000,10),
  read_diff_histo_bin_width = 10
  )
GAL_count_diff_df <- compare_bcwithqc_align_UMI_tools(
  data_type = "reads",
  cfg = cfg,
  bcwithqc_dir = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_bcwithqc/bcwithqc_output",
  align_UMI_tools_read_folder = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_align/mapped",
  read_diff_histo_x_lim = c(-1000,10),
  read_diff_histo_bin_width = 10
  )
# test_df <- compare_bcwithqc_align_UMI_tools(
#   data_type = "reads",
#   cfg = cfg,
#   bcwithqc_dir = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_bcwithqc/bcwithqc_output",
#   align_UMI_tools_read_folder = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_align/mapped",
#   read_diff_histo_x_lim = c(-100,5),
#   read_diff_histo_bin_width = 1
#   )
```

``` {r compare_bcwithqc_vs_align_UMI_tools, eval=FALSE}

# CURRENTLY I HAVE THESE SET TO THE OLD FILES!
xlsx_files <- c(
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_align/results/align_UMI_tools_reads_control_median_RI_comb_sample_comb_none_mapping_results.xlsx",
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_bcwithqc/results/mapping_results.xlsx",
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_align/results/align_UMI_tools_reads_control_median_RI_comb_sample_comb_none_mapping_results.xlsx",
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_bcwithqc/results/mapping_results.xlsx",
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_align/results/align_UMI_tools_reads_control_median_RI_comb_sample_comb_none_mapping_results.xlsx",
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_bcwithqc/results/mapping_results.xlsx"
)

names <- c(
  "DCA align",
  "DCA bcwithqc",
  "PA align",  
  "PA bcwithqc",
  "GAL align",  
  "GAL bcwithqc"
)

output_dir <- "/g/steinmetz/link/Presentations/"

plots <- plot_mapping_qc_boxes(
  xlsx_paths = xlsx_files,
  sample_names = names,
  output_dir = output_dir
)


```

``` {r compare_hits_bcwithqc_vs_align_UMI_tools, eval=TRUE}

xlsx_files <- c(
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_align/results/Hits_explorative_align_UMI_tools_reads_control_median_RI_comb_sample_comb_none.xlsx",
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_bcwithqc/results/Hits_explorative.xlsx",
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_align/results/Hits_explorative_align_UMI_tools_reads_control_median_RI_comb_sample_comb_none.xlsx",
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_bcwithqc/results/Hits_explorative.xlsx",
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_align/results/Hits_explorative_align_UMI_tools_reads_control_median_RI_comb_sample_comb_none.xlsx",
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_bcwithqc/results/Hits_explorative.xlsx"
)

names <- c(
  "DCA align",
  "DCA bcwithqc",
  "PA align",
  "PA bcwithqc",
  "GAL align",
  "GAL bcwithqc"
)

output_dir <- "/g/steinmetz/link/Presentations/John_Update_2026_06/explorative_NEW"

hit_compare_results <- compare_hit_xlsx_pairs(
  xlsx_paths = xlsx_files,
  sample_names = names,
  output_dir = output_dir,
  rank_method = "spearman"
)

``` 

``` {r qq_rectangle_plot, eval=TRUE}

pa_align_mean_rds_fpath <- 
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/Mean_results/align/mean_results_PA.rds"
dca_align_mean_rds_fpath <- 
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/Mean_results/align/mean_results_DCA.rds"
gal_align_mean_rds_fpath <- 
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/Mean_results/align/mean_results_GALNAC.rds"
pa_bcwithqc_mean_rds_fpath <- 
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/Mean_results/bcwithqc/PA_mean_hits_bcwithqc.rds"
dca_bcwithqc_mean_rds_fpath <- 
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/Mean_results/bcwithqc/DCA_mean_hits_bcwithqc.rds"
gal_bcwithqc_mean_rds_fpath <- 
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/Mean_results/bcwithqc/GAL_mean_hits_bcwithqc.rds"
nico_dir <- "/g/steinmetz/link/sharing/nico/qq_rectangle_plots"


p <- qq_rectangle_plot(
  align_rds_fpath = pa_align_mean_rds_fpath,
  bcwithqc_rds_fpath = pa_bcwithqc_mean_rds_fpath,
  title = "PA bcwithqc vs align rank comparison",
  FDR_thresh = 0.05,
  rank_gamma = 20, # rank_gamma = 1 is linear scalling
  grey_frac = 0.2,
  save_fpath = file.path(nico_dir, "PA_squish.png")
)
print(p)
p <- qq_rectangle_plot(
  align_rds_fpath = dca_align_mean_rds_fpath,
  bcwithqc_rds_fpath = dca_bcwithqc_mean_rds_fpath,
  title = "DCA bcwithqc vs align rank comparison",
  FDR_thresh = 0.05,
  rank_gamma = 20, # rank_gamma = 1 is linear scalling
  grey_frac = 0.2,
  save_fpath = file.path(nico_dir, "DCA_squish.png")
)
print(p)
p <- qq_rectangle_plot(
  align_rds_fpath = gal_align_mean_rds_fpath,
  bcwithqc_rds_fpath = gal_bcwithqc_mean_rds_fpath,
  title = "GalNAc bcwithqc vs align rank comparison",
  FDR_thresh = 0.05,
  rank_gamma = 20, # rank_gamma = 1 is linear scalling
  grey_frac = 0.2,
  save_fpath = file.path(nico_dir, "GAL_squish.png")
)
print(p)
combined_delta_pa <- make_combined_delta_df(
  align_rds_fpath = pa_align_mean_rds_fpath,
  bcwithqc_rds_fpath = pa_bcwithqc_mean_rds_fpath,
  FDR_thresh = 0.05
)
saveRDS(combined_delta_pa, file.path(nico_dir, "PA_changed_hits.rds"))
combined_delta_dca <- make_combined_delta_df(
  align_rds_fpath = dca_align_mean_rds_fpath,
  bcwithqc_rds_fpath = dca_bcwithqc_mean_rds_fpath,
  FDR_thresh = 0.05
)
saveRDS(combined_delta_dca, file.path(nico_dir, "DCA_changed_hits.rds"))
combined_delta_gal <- make_combined_delta_df(
  align_rds_fpath = gal_align_mean_rds_fpath,
  bcwithqc_rds_fpath = gal_bcwithqc_mean_rds_fpath,
  FDR_thresh = 0.05
)
saveRDS(combined_delta_gal, file.path(nico_dir, "GAL_changed_hits.rds"))

``` 


```{r further_analysis_read_count_diff, eval=TRUE}
df <- PA_count_diff_df
df <- df %>% 
  left_join(cfg$merged_sgRNA_df %>% 
              select(-sublib), by = c("sgRNA" = "sgrna_id")) %>% 
  select(-c(count, Gene, align_seq))
df <- df %>% 
  filter(condition == "input") %>% 
  filter(exp == "L1_1")


df <- df %>% 
  filter(count_bcwithqc == 1)

library(Matrix)

mat <- readMM(gzfile("/home/link/John_UMIs/bcwithqc/tests/gDNA_single_end_no_STAR/raw_reads_bc_matrix/matrix.mtx.gz"))
mat <- as(mat, "dgCMatrix")

barcodes <- readLines(gzfile("/home/link/John_UMIs/bcwithqc/tests/gDNA_single_end_no_STAR/raw_reads_bc_matrix/barcodes.tsv.gz"))

if (ncol(mat) == length(barcodes)) {
  colnames(mat) <- barcodes
} else if (nrow(mat) == length(barcodes)) {
  rownames(mat) <- barcodes
} else {
  stop("Barcode number does not match matrix rows or columns.")
}

if (is.null(rownames(mat))) {
  rownames(mat) <- paste0("feature_", seq_len(nrow(mat)))
}

df_3 <- data.frame(
  barcode = colnames(mat),
  value = as.numeric(mat[1, ])
)
```

## Functions that add more information to the results, like the gene symbol

```{r check_MAUDE_results, eval=TRUE}

```

## Create QQplots for the p-values. 
Lots of debris from old runs in here. Use with caution

```{r QQplot_functions, eval=TRUE}
plot_qq_maude <- function(maude_gene_stats, data_name, mark_gene = "AHSA1") {
  # Check that required columns exist
  if (!"p.value" %in% names(maude_gene_stats) || !"symbol" %in% names(maude_gene_stats)) {
    stop("Input data frame must contain 'p.value' and 'symbol' columns.")
  }
  
  # Sort the observed p-values and keep corresponding symbols
  sorted_stats <- maude_gene_stats[order(maude_gene_stats$p.value), ]
  # observed_pvalues <- p.adjust(sorted_stats$p.value, method = "hochberg", n = length(sorted_stats$p.value))
  observed_pvalues <- sorted_stats$p.value
  gene_symbols <- sorted_stats$symbol
  
  # Expected p-values adjusted to observed max
  max_p <- max(observed_pvalues)
  expected_pvalues <- (1:length(observed_pvalues)) / (length(observed_pvalues) + 1) * max_p
  
  # Transform to -log10
  log_observed <- -log10(observed_pvalues)
  log_expected <- -log10(expected_pvalues)
  
  # Identify the point to mark
  mark_index <- which(gene_symbols == mark_gene)
  
  # Plot Q-Q plot
  plot(log_expected, log_observed,
       main = paste("p-values Q-Q plot for", data_name),
       xlab = "Expected -log(p-value)",
       ylab = "Observed -log(p-value)",
       pch = 20, col = "blue",
       xlim = c(0, 5), ylim = c(0, 5))
  
  # Reference line
  abline(0, 1, col = "red", lty = 2)
  
  # Highlight the specific gene
  if (length(mark_index) > 0) {
    points(log_expected[mark_index], log_observed[mark_index],
           col = "darkorange", pch = 17, cex = 1.5)
    text(log_expected[mark_index], log_observed[mark_index],
         labels = mark_gene, pos = 4, offset = 0.5, col = "darkorange")
  }
}

get_top_N <- function(df, N = 1000, lookup_df = merged_sgRNA_df){
  N_half <- round(N / 2) 
  df_N <- bind_rows(
    df %>% arrange(significanceZ) %>% head(N_half),      # lowest Z
    df %>% arrange(desc(significanceZ)) %>% head(N_half) # highest Z
  )
  return(df_N)
}

get_top_N_intersect <- function(df_1,df_2,N = 1000, lookup_df = merged_sgRNA_df){
  
  N_half <- round(N / 2)  
  
  # Get top and bottom N/2 from df_1
  df1_extreme <- bind_rows(
    df_1 %>% arrange(significanceZ) %>% head(N_half),      # lowest Z
    df_1 %>% arrange(desc(significanceZ)) %>% head(N_half) # highest Z
  )
  
  # Same for df_2
  df2_extreme <- bind_rows(
    df_2 %>% arrange(significanceZ) %>% head(N_half),
    df_2 %>% arrange(desc(significanceZ)) %>% head(N_half)
  )
  # Find common entrez values
  common_entrez <- base::intersect(df1_extreme$entrez, df2_extreme$entrez)
  
  # Filter both for common entrez
  df1_common <- df1_extreme %>% filter(entrez %in% common_entrez)
  df2_common <- df2_extreme %>% filter(entrez %in% common_entrez)
  
  # Create combined results
  results_df <- bind_rows(df1_common, df2_common) %>%
    distinct(entrez, .keep_all = TRUE)  # one row per entrez, from first found in either df
  
  # Join with lookup — only first match per entrez
  lookup_first <- lookup_df %>%
    group_by(entrez) %>%
    dplyr::slice(1) %>%  # keep only the first match
    ungroup() %>%
    select(entrez, sgrna_id, symbol) %>% 
    mutate(entrez = as.character(entrez))
  
  # Add to results
  results_df <- results_df %>%
    left_join(lookup_first, by = "entrez")
  
  return(results_df)
}

compare_versions <- function(base_name, file_path = "/home/link/NB_EXP030/MAUDE_gene_stats") {
  # Create a vector of suffixes
  suffixes <- c("", "_RI", "_D0", "_RI_D0")
  
  # Read the files based on combinations of base_name and suffixes
  files <- lapply(suffixes, function(suffix) {
    file <- paste0(file_path, "_", base_name, suffix, ".rds")
    readRDS(file)
  })
  
  # Assign meaningful names to each version (e.g., "align_UMI_tools_reads", "bcwithqc_reads", etc.)
  names(files) <- paste0(base_name, suffixes)
  
  # Perform the comparisons (you can adjust this logic depending on your requirements)
  comparisons <- list()
  for (i in 1:length(files)) {
    for (j in i:length(files)) {
      if (i != j) {
        comparison_name <- paste(names(files)[i], "vs", names(files)[j])
        comparisons[[comparison_name]] <- get_top_N_intersect(files[[i]], files[[j]])
      }
    }
  }
  
  return(comparisons)
}


compare_versions <- function(file_path = get("rds_output_folder", envir = .GlobalEnv),
                             base_names = c("align_UMI_tools_reads", "bcwithqc_reads", "align_UMI_tools_umis", "bcwithqc_rf_umis")) {

  # Create an empty list to store all comparisons
  comparisons <- list()

  # Suffixes to consider
  suffixes <- c("", "_RI", "_D0", "_RI_D0")
  
  AHSA1_position_finder <- function(file, name){
    # Sort the dataframe by 'significanceZ' in ascending order
    sorted_df <- file %>%
      arrange(significanceZ)  # Assuming 'significanceZ' is the column name
  
    # Find the row where the 'entrez' value is 10598
    target_row <- sorted_df %>% filter(entrez == 10598)
    
    if (nrow(target_row) > 0) {
      # Get the position (row number) and significanceZ value
      position <- which(sorted_df$entrez == 10598)
      significanceZ_value <- target_row$significanceZ
      
      # Print the result
      cat(paste("AHSA1 position in", name, "is", position, "with significanceZ value of", significanceZ_value, "\n"))
    } else {
      cat(paste("AHSA1 (entrez 10598) not found in", name, "\n"))
    }
  }
  
  name_list <- list()
  # Comparison 1: Compare all versions of each base name against each other
  for (base_name in base_names) {
    # Read files based on the base_name and suffixes
    files <- lapply(suffixes, function(suffix) {
      file <- paste0(file_path, "_", base_name, suffix, ".rds")
      if (file.exists(file)) {
        name_list <<- c(name_list, paste0(base_name, suffix))  # Use <<- to modify the global name_list
        return(readRDS(file))
      } else {
        cat(paste("File not found, skipping:", file, "\n"))
        return(NULL)  # Return NULL if file doesn't exist
      }
    })
    
    # Remove any NULL elements (files that were missing)
    files <- files[!sapply(files, is.null)]
    names(files) <- name_list

    # Perform comparisons between all versions of the base name
    for (i in 1:length(files)) {
      for (j in i:length(files)) {
        cat(paste("------------------------------------------------","\n"))
        if (i != j) {
          cat(paste("------------------------------------------------","\n"))
          comparison_name <- paste(names(files)[i], "vs", names(files)[j])
          intersect_df <- get_top_N_intersect(files[[i]], files[[j]], N = 100)
          cat(paste(comparison_name, "Number of intersects:",nrow(intersect_df),"\n"))
          AHSA1_position_finder(files[[i]], names(files)[i])
          AHSA1_position_finder(files[[j]], names(files)[j])
          comparisons[[paste(base_name, comparison_name)]] <- intersect_df
        }
      }
    }
  }
  
  # Comparison 2: Compare reads and umis between bcwithqc and align_UMI_tools
  # Define pairs for align_UMI_tools vs. bcwithqc comparisons for "reads" and "umis"
  bcwithqc_align_UMI_tools_comparisons <- list(
    c("align_UMI_tools_reads", "bcwithqc_reads"),
    c("align_UMI_tools_umis", "bcwithqc_rf_umis")
  )
  
  for (pair in bcwithqc_align_UMI_tools_comparisons) {
    for (suffix in suffixes) {
      cat(paste("------------------------------------------------","\n"))
      # Construct filenames for align_UMI_tools and bcwithqc based on base names and suffixes
      align_UMI_tools_file <- paste0(file_path, "_", pair[1], suffix, ".rds")
      bcwithqc_file <- paste0(file_path, "_", pair[2], suffix, ".rds")
      
      # Check if the files exist
      if (file.exists(align_UMI_tools_file) & file.exists(bcwithqc_file)) {
        # Read the files
        align_UMI_tools_data <- readRDS(align_UMI_tools_file)
        bcwithqc_data <- readRDS(bcwithqc_file)
        
        # Perform the comparison between align_UMI_tools and bcwithqc for each suffix combination
        comparison_name <- paste(pair[1], "vs", pair[2], "(", suffix, ")")
        intersect_df <- get_top_N_intersect(align_UMI_tools_data, bcwithqc_data, N = 100)
        cat(paste(comparison_name, "Number of intersects:",nrow(intersect_df),"\n"))
        comparisons[[paste(pair[1], pair[2], comparison_name)]] <- intersect_df
      } else {
        cat(paste("One or both files not found, skipping:", align_UMI_tools_file, "and", bcwithqc_file, "\n"))
      }
    }
  }

  return(comparisons)
}
plot_all_qq_plots <- function(dir_path) {
  # List all files in the directory ending with ".rds"
  rds_files <- list.files(dir_path, pattern = "\\.rds$", full.names = TRUE)

  # Filter files that start with "MAUDE_guide_stats_"
  matching_files <- rds_files[grepl("MAUDE_guide_stats_", basename(rds_files))]

  # Loop over each matching file
  for (file in matching_files) {
    # Extract the info_suffix
    filename <- basename(file)
    info_suffix <- sub("^MAUDE_guide_stats_", "", sub("\\.rds$", "", filename))

    # Run the commands
    Hits_current_settings <- add_info_wrapper(info_suffix)
    plot_qq_maude(Hits_current_settings, info_suffix)
  }
}


generate_qq_plots <- function(base_name,
                              file_path = get("rds_output_folder", envir = .GlobalEnv),
                              suffixes = c("", "_RI", "_D0", "_RI_D0")) {
  
  file_path <- paste0(file_path, "MAUDE_gene_stats")
  # Use invisible() to suppress unwanted NULL outputs
  invisible(lapply(suffixes, function(suffix) {
    # Create the file path for each suffix
    file <- paste0(file_path, "_", base_name, suffix, ".rds")
    
    # Check if the file exists before attempting to read it
    if (file.exists(file)) {
      maude_gene_stats <- readRDS(file)
      plot_qq_maude(maude_gene_stats, paste(base_name, suffix))
    } else {
      # If the file doesn't exist, print a message and skip this iteration
      cat(paste("File not found, skipping:", file, "\n"))
    }
  }))
}
compare_Hits <- function(df1, df2, top_N = 50) {
  
  # Sort both dataframes by FDR (low to high)
  df1_sorted <- df1[order(df1$FDR), ]
  df2_sorted <- df2[order(df2$FDR), ]
  
  # Get the top N entries based on FDR
  df1_top <- df1_sorted[1:top_N, ]
  df2_top <- df2_sorted[1:top_N, ]
  
  # 1. Overlap among top_N (by FDR)
  overlap <- merge(df1_top, df2_top, by = "entrez", all = FALSE)
  
  # 2. Dropouts among top_N (by FDR)
  dropouts <- base::setdiff(df1_top$entrez, df2_top$entrez)
  df1_dropouts <- df1_top[df1_top$entrez %in% dropouts, ]
  
  # 3. Statistics dataframe
  # Number of overlaps
  num_overlaps <- nrow(overlap)
  
  # Number of dropouts
  num_dropouts <- nrow(df1_dropouts)
  
  # Calculate ranks for FDR in both dataframes
  df1_sorted$rank_FDR <- rank(df1_sorted$FDR)
  df2_sorted$rank_FDR <- rank(df2_sorted$FDR)
  
  # Merge the two sorted dataframes by 'entrez' to match pairs
  merged_df <- merge(df1_sorted[, c("entrez", "rank_FDR")], 
                     df2_sorted[, c("entrez", "rank_FDR")], 
                     by = "entrez", 
                     suffixes = c("_df1", "_df2"))
  
  # 4. Correlation between ranks in top N (paired test)
  rank_correlation_topN <- cor(merged_df$rank_FDR_df1[1:top_N], merged_df$rank_FDR_df2[1:top_N], method = "pearson")
  
  # 5. Correlation between ranks for all entries (paired test)
  rank_correlation_all <- cor(merged_df$rank_FDR_df1, merged_df$rank_FDR_df2, method = "pearson")
  
  # Split significanceZ into > 0 and <= 0
  df1_top_posZ <- df1_top[df1_top$significanceZ > 0, ]
  df2_top_posZ <- df2_top[df2_top$significanceZ > 0, ]
  
  df1_top_negZ <- df1_top[df1_top$significanceZ <= 0, ]
  df2_top_negZ <- df2_top[df2_top$significanceZ <= 0, ]
  
  # Median and mean changes in significance Z and FDR for significanceZ > 0
  median_change_Z_pos <- median(df1_top_posZ$significanceZ - df2_top_posZ$significanceZ)
  mean_change_Z_pos <- mean(df1_top_posZ$significanceZ - df2_top_posZ$significanceZ)
  
  # Median and mean changes in significance Z and FDR for significanceZ <= 0
  median_change_Z_neg <- median(df1_top_negZ$significanceZ - df2_top_negZ$significanceZ)
  mean_change_Z_neg <- mean(df1_top_negZ$significanceZ - df2_top_negZ$significanceZ)
  
  # Median and mean changes in FDR
  median_change_FDR <- median(df1_top$FDR - df2_top$FDR)
  mean_change_FDR <- mean(df1_top$FDR - df2_top$FDR)
  
  # Create the statistics dataframe
  stats <- data.frame(
    Statistic = c(paste0("Number of overlaps (top ",top_N,")"),paste0("Number of dropouts (top ",top_N,")"), 
                  paste0("Correlation between ranks (top ",top_N,")"), "Correlation between ranks (all)", 
                  "Median change in significance Z+", "Mean change in significance Z+", 
                  "Median change in significance Z-", "Mean change in significance Z-", 
                  "Median change in FDR", "Mean change in FDR"),
    Value = c(num_overlaps, num_dropouts, rank_correlation_topN, rank_correlation_all, 
              median_change_Z_pos, mean_change_Z_pos, median_change_Z_neg, mean_change_Z_neg, 
              median_change_FDR, mean_change_FDR)
  )
  
  # Return the list of dataframes
  return(list(
    Overlap = overlap,
    Dropouts = df1_dropouts,
    Statistics = stats
  ))
}

# Create a smaller version of compare_versions for the specific four dataframes
compare_methods <- function(N = 100) {
  # List of the four dataframes
  df_list <- list(
    "Hits_align_UMI_tools_umis" = Hits_align_UMI_tools_umis,
    "Hits_align_UMI_tools_umis_RI" = Hits_align_UMI_tools_umis_RI,
    "Hits_align_UMI_tools_umis_sum_RI" = Hits_align_UMI_tools_umis_sum_RI,
    "Hits_align_UMI_tools_umis_rep_RI" = Hits_align_UMI_tools_umis_rep_RI
  )
  
  # Helper function to calculate AHSA1 position
  AHSA1_position_finder <- function(df, name) {
    sorted_df <- df %>%
      arrange(significanceZ)  # Sort by significanceZ (assuming column exists)
    
    # Find the row where the 'entrez' value is 10598
    target_row <- sorted_df %>% filter(entrez == 10598)
    
    if (nrow(target_row) > 0) {
      position <- which(sorted_df$entrez == 10598)
      significanceZ_value <- target_row$significanceZ
      cat(paste("AHSA1 position in", name, "is", position, "with significanceZ value of", significanceZ_value, "\n"))
    } else {
      cat(paste("AHSA1 (entrez 10598) not found in", name, "\n"))
    }
  }
  
  # Compare all combinations of the dataframes
  comparisons <- list()
  
  df_names <- names(df_list)
  
  # Loop through all possible pairs of dataframes
  for (i in 1:(length(df_names) - 1)) {
    cat(paste("------------------------------------------------","\n"))
    for (j in (i + 1):length(df_names)) {
      cat(paste("------------------------------------------------","\n"))
      name_i <- df_names[i]
      name_j <- df_names[j]
      
      # Get the dataframes
      df_i <- df_list[[name_i]]
      df_j <- df_list[[name_j]]
      
      # Calculate the intersection
      intersect_df <- get_top_N_intersect(df_i, df_j, N = N)
      n_intersect <- nrow(intersect_df)
      
      # Print intersection and AHSA1 positions
      cat(paste(name_i, "vs", name_j, "Number of intersects:", n_intersect, "\n"))
      AHSA1_position_finder(df_i, name_i)
      AHSA1_position_finder(df_j, name_j)
      
      # Store the results
      comparisons[[paste(name_i, "vs", name_j)]] <- list(
        intersect = intersect_df,
        n_intersect = n_intersect
      )
    }
  }
  
  return(comparisons)
}

```

## Create QQplots for the p-values. 
Lots of debris from old runs in here. Use with caution

```{r QQplot, eval=FALSE}
#===============================================================================
# Batch generate ALL qq plots
# add suffixes <- c(list_of_suffixes_without_.rds); example <- c("", "_RI")
#===============================================================================
plot_all_qq_plots(rds_output_folder)
#===============================================================================
# Batch generate specific qq plots
# add suffixes <- c(list_of_suffixes_without_.rds); example <- c("", "_RI")
#===============================================================================
generate_qq_plots("align_UMI_tools_reads")
generate_qq_plots("align_UMI_tools_umis")
generate_qq_plots("bcwithqc_reads")
generate_qq_plots("bcwithqc_rf_umis")

#===============================================================================
# Create Hit lists
#===============================================================================
# use my_Hits <- add_info_wrapper("my_file_suffix")
Hits_current_settings <- add_info_wrapper(file_info_suffix)
plot_qq_maude(Hits_current_settings, "Sublibraries as replicates combined")
plot_qq_maude(export_df, "Sublibraries as replicates combined")
Hits_sublib <- add_info_wrapper("align_UMI_tools_umis_rep_sublib_control_median_RI")
Hits_align_UMI_tools_umis_control_median_RI <- add_info_wrapper("align_UMI_tools_umis_control_median_RI")
Hits_align_UMI_tools_umis_sum_control_median_RI <- add_info_wrapper("align_UMI_tools_umis_sum_control_median_RI")
Hits_sublib_combined <- add_info_wrapper("align_UMI_tools_umis_rep_sublib_control_median_RI_0")

#===============================================================================
# Intersect 2 different Hit lists
#===============================================================================
top_1000_umis_intersect <- get_top_N_intersect(Hits_default,Hits_sublib_combined, N=1000)
top_100_umis_intersect <- get_top_N_intersect(Hits_default,Hits_sublib_combined, N=100)
top_100 <- get_top_N_intersect(Hits_current_settings,Hits_default, N=100)
top_100_2 <- get_top_N_intersect(Hits_current_settings,Hits_sum, N=100)



top_50_umis_intersect <- get_top_N_intersect(Hits_align_UMI_tools_umis_RI,Hits_align_UMI_tools_umis_sum_RI, N=50)
top_20_umis_intersect <- get_top_N_intersect(Hits_align_UMI_tools_umis_RI,Hits_align_UMI_tools_umis_sum_RI, N=20)
top_50_umis_intersect_sum_rep <- get_top_N_intersect(Hits_align_UMI_tools_reads_sum_RI,Hits_align_UMI_tools_reads_rep_RI_rf, N=50)
top_10_umis_intersect___sum <- get_top_N_intersect(Hits_align_UMI_tools_umis_sum_RI,Hits_align_UMI_tools_umis_RI, N=10)
top_15_umis_intersect_control_median <- get_top_N_intersect(Hits_align_UMI_tools_umis_sum_control_median_RI,Hits_align_UMI_tools_umis_control_median_RI, N=100)
top_10_umis_intersect_both <- get_top_N_intersect(top_15_umis_intersect_control_median,top_10_umis_intersect___sum, N=10)
top_10_umis_sum_control_median <- get_top_N_intersect(Hits_align_UMI_tools_umis_sum_control_median_RI,Hits_align_UMI_tools_umis_sum_control_median_RI, N=10)
saveRDS(get_top_N(Hits_align_UMI_tools_umis_sum_RI, N=100), "/g/steinmetz/link/Amplicon_barcode_analysis/NB_EXP030/heatmaps/top_100.rds")
saveRDS(get_top_N(Hits_align_UMI_tools_umis_sum_RI, N=50), "/g/steinmetz/link/Amplicon_barcode_analysis/NB_EXP030/heatmaps/top_50.rds")
saveRDS(get_top_N(Hits_align_UMI_tools_umis_sum_RI,N=20), "/g/steinmetz/link/Amplicon_barcode_analysis/NB_EXP030/heatmaps/top_20.rds")
#===============================================================================
# Make individual qq-plots for the p-values of the hits. 
#===============================================================================
plot_qq_maude(Hits_align_UMI_tools_umis_RI, "UMIs (one replicate method)")
plot_qq_maude(Hits_align_UMI_tools_umis_sum_RI, "UMIs (sum method)")
plot_qq_maude(Hits_align_UMI_tools_reads_RI, "reads (one replicate method)")
plot_qq_maude(Hits_align_UMI_tools_reads_sum_RI, "reads (sum method)")

plot_qq_maude(Hits_align_UMI_tools_umis_control_median_RI, "UMIs (one replicate method)(normalized)")
plot_qq_maude(Hits_align_UMI_tools_umis_sum_control_median_RI, "UMIs (sum method)(normalized)")
plot_qq_maude(Hits_align_UMI_tools_reads_control_median_RI, "reads (one replicate method)(normalized)")
plot_qq_maude(Hits_align_UMI_tools_reads_sum_control_median_RI, "reads (sum method)(normalized)")

plot_qq_maude(Hits_align_UMI_tools_reads, "Reads (no IR no RF)")
plot_qq_maude(Hits_bcwithqc_reads, "bcwithqc Reads (no IR no RF)")
plot_qq_maude(Hits_align_UMI_tools_umis, "UMIs (no IR)")
plot_qq_maude(Hits_bcwithqc_umis, "bcwithqc UMIs (no IR no RF)")
plot_qq_maude(Hits_bcwithqc_rf_umis, "bcwithqc_rf UMIs (no IR)")

plot_qq_maude(Hits_align_UMI_tools_reads_RI, "Reads (no IR)")
plot_qq_maude(Hits_bcwithqc_reads_RI, "bcwithqc Reads (no RF)")
plot_qq_maude(Hits_align_UMI_tools_reads_RI_rf, "Reads")
# plot_qq_maude(Hits_bcwithqc_reads_RI_rf, "bcwithqc Reads")

plot_qq_maude(Hits_align_UMI_tools_umis_RI, "UMIs")
plot_qq_maude(Hits_bcwithqc_rf_umis_RI, "bcwithqc UMIs")

plot_qq_maude(Hits_align_UMI_tools_reads_sum_RI_rf, "Reads replicate summation")
plot_qq_maude(Hits_bcwithqc_reads_sum_RI_rf, "bcwithqc Reads replicate summation")
plot_qq_maude(Hits_align_UMI_tools_umis_sum_RI, "UMIs replicate summation")
plot_qq_maude(Hits_bcwithqc_rf_umis_sum_RI, "bcwithqc UMIs replicate summation")

plot_qq_maude(Hits_align_UMI_tools_reads_rep_RI_rf, "Reads multiple replicates")
plot_qq_maude(Hits_bcwithqc_reads_rep_RI_rf, "bcwithqc Reads multiple replicates")
plot_qq_maude(Hits_align_UMI_tools_umis_rep_RI, "UMIs multiple replicates")
plot_qq_maude(Hits_bcwithqc_rf_umis_rep_RI, "bcwithqc UMIs multiple replicates")





# Call the function to perform the comparisons
results_methods <- compare_methods(N = 50)



# maude_gene_stats_align_UMI_tools <- readRDS(paste0("/home/link/NB_EXP030/MAUDE_gene_stats", "_align_UMI_tools_reads.rds"))
# maude_gene_stats_bcwithqc <- readRDS(paste0("/home/link/NB_EXP030/MAUDE_gene_stats", "_bcwithqc_reads.rds"))
# 
# top_1000_reads_intersect <- get_top_N_intersect(maude_gene_stats_align_UMI_tools,maude_gene_stats_bcwithqc)
# top_100_reads_intersect <- get_top_N_intersect(maude_gene_stats_align_UMI_tools,maude_gene_stats_bcwithqc, N=100)

```

``` {r compare_mapping_results, eval=FALSE}
pa_mapping_file_path <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA/results/align_UMI_tools_reads_control_median_RI_comb_sample_mapping_results.xlsx"
dca_mapping_file_path <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA/results/align_UMI_tools_reads_control_median_RI_comb_sample_mapping_results.xlsx"
gal_mapping_file_path <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC/results/align_UMI_tools_reads_control_median_RI_comb_sample_mapping_results.xlsx"
library(readxl)
map_pa_df  <- read_excel(pa_mapping_file_path)
map_dca_df <- read_excel(dca_mapping_file_path)
map_gal_df <- read_excel(gal_mapping_file_path)

# helper to convert "10.12%" -> 10.12
convert_percent <- function(x) {
  as.numeric(str_remove(x, "%"))
}

# combine into one plotting dataframe
plot_df <- bind_rows(
  data.frame(group = "PA",     wrong_perc = convert_percent(map_pa_df$wrong_perc)),
  data.frame(group = "DCA",    wrong_perc = convert_percent(map_dca_df$wrong_perc)),
  data.frame(group = "GalNAc", wrong_perc = convert_percent(map_gal_df$wrong_perc))
)

# optional: remove missing values
plot_df <- plot_df %>% filter(!is.na(wrong_perc))

# boxplot
ggplot(plot_df, aes(x = group, y = wrong_perc, fill = group)) +
  geom_boxplot() +
  scale_fill_manual(values = c(
    "PA" = "#4DAF4A",
    "DCA" = "#377EB8",
    "GalNAc" = "#E41A1C"
  )) +
  labs(
    x = NULL,
    y = "Percentage of reads mapped to wrong sublibrary (%)",
    title = "Reads mapped to the wrong sublibrary"
  ) +
  theme_bw() +
  theme(legend.position = "none")
``` 


``` {r Calculated_replicated_analysis, eval=False}
if(first_time){
  high_confidence_hits <- automate_calc_replicate_comparison(FDR_threshold = 0.05,
                                                           hits_in_X_reps = 16,
                                                           correlation_heatmap = FALSE,
                                                           venn_diagram = FALSE)
  explorative_hits <- automate_calc_replicate_comparison(FDR_threshold = 0.05,
                                                       hits_in_X_reps = 10,
                                                       correlation_heatmap = TRUE,
                                                       venn_diagram = FALSE)

  excel_file_path <- sub("\\.rds$", ".xlsx", paste0(results_output_folder, "/", "Hits_high_confidence", file_suffix))
  write_xlsx(high_confidence_hits$Hits_in_X_df, excel_file_path)
  saveRDS(high_confidence_hits$Hits_in_X_df, get_file_path(rds_output_folder, paste0("high_confidence_hits","_",file_info_suffix,".rds")))
  
  excel_file_path <- sub("\\.rds$", ".xlsx", paste0(results_output_folder, "/", "Hits_explorative", file_suffix))
  write_xlsx(explorative_hits$Hits_in_X_df, excel_file_path)
  saveRDS(explorative_hits$Hits_in_X_df, get_file_path(rds_output_folder, paste0("explorative_hits","_",file_info_suffix,".rds")))
  
  
  PA_means <- automate_calc_replicate_means(
    cfg,
    file_suffix = ".rds",
    rds_output_folder="/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_bcwithqc/rds")
  DCA_means <- automate_calc_replicate_means(
    cfg,
    file_suffix = ".rds",
    rds_output_folder="/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_bcwithqc/rds")
  GALNAC_means <- automate_calc_replicate_means(
    cfg,
    file_suffix = ".rds",
    rds_output_folder="/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_bcwithqc/rds")
  
  dual_rep_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/"
  
  saveRDS(PA_means,"/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_bcwithqc/rds/PA_mean_hits_bcwithqc.rds")
  saveRDS(DCA_means,"/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_bcwithqc/rds/DCA_mean_hits_bcwithqc.rds")
  saveRDS(GALNAC_means,"/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_bcwithqc/rds/GAL_mean_hits_bcwithqc.rds")
  
  write_xlsx(PA_means,"/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_bcwithqc/rds/PA_mean_hits_bcwithqc.xlsx")
  write_xlsx(DCA_means,"/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_bcwithqc/rds/DCA_mean_hits_bcwithqc.xlsx")
  write_xlsx(GALNAC_means,"/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_bcwithqc/rds/GAL_mean_hits_bcwithqc.xlsx")
  
  mean_results_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/Mean_results_bcwithqc"
  
  saveRDS(PA_means, file.path(mean_results_dir, "PA_mean_hits_bcwithqc.rds"))
  saveRDS(DCA_means, file.path(mean_results_dir, "DCA_mean_hits_bcwithqc.rds"))
  saveRDS(GALNAC_means, file.path(mean_results_dir, "GAL_mean_hits_bcwithqc.rds"))
  
  write_xlsx(PA_means, file.path(mean_results_dir, "PA_mean_hits_bcwithqc.xlsx"))
  write_xlsx(DCA_means, file.path(mean_results_dir, "DCA_mean_hits_bcwithqc.xlsx"))
  write_xlsx(GALNAC_means, file.path(mean_results_dir, "GAL_mean_hits_bcwithqc.xlsx"))
} else {
  high_confidence_hits <- readRDS(get_file_path(rds_output_folder, paste0("high_confidence_hits","_",file_info_suffix,".rds")))
  explorative_hits <- readRDS(get_file_path(rds_output_folder, paste0("explorative_hits","_",file_info_suffix,".rds")))
  
  PA_means <- readRDS(file.path(mean_results_dir, "mean_results_PA.rds"))
  DCA_means <- readRDS(file.path(mean_results_dir, "mean_results_DCA.rds"))
  GALNAC_means <- readRDS(file.path(mean_results_dir, "mean_results_GALNAC.rds"))
}

```
## Compare Sublibraries 
For this to function You need to run data_analysis_with_MAUDE multiple times,
once without excluding sublibraries, and then at least once more with
one or more sublibraries excluded using the skip_list_sublib setting.
``` {r compare_sublibraries_analysis, eval=False}
test <- automate_sublibrary_comparison(FDR_threshold = 0.05)
```

``` {r compare_subsamples_analysis, eval=False}
if (FALSE){
  # run this once to move subsamples from the target dir, to the rds dir
  # of the current output folder
  move_subsamples("/g/steinmetz/link/Amplicon_barcode_analysis/PA_subsampeling/1/HepG2_dual_rep_PA_subsample", subsample_prefix = "A")
  move_subsamples("/g/steinmetz/link/Amplicon_barcode_analysis/PA_subsampeling/2/HepG2_dual_rep_PA_subsample", subsample_prefix = "B")
  move_subsamples("/g/steinmetz/link/Amplicon_barcode_analysis/PA_subsampeling/3/HepG2_dual_rep_PA_subsample", subsample_prefix = "C")
  move_subsamples("/g/steinmetz/link/Amplicon_barcode_analysis/PA_subsampeling/4/HepG2_dual_rep_PA_subsample", subsample_prefix = "D")

 
automate_subsample_comparison(FDR_threshold = 0.05)
```

``` {r get_validation_list, eval=False}
source(file.path(project_root_dir, "functions", "zzz_source_all.R"))
dual_rep_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/"

if (first_time) {
  all_conjugates_overlap_df <- read.csv2(
    file.path(dual_rep_dir,"Explorative_Hits_Overlap_between_conjugates.csv"),
    header = TRUE
  ) 
  all_conjugates_overlap_df <- calculate_category_based_Z_score(all_conjugates_overlap_df)
  all_conjugates_overlap_classified <- assign_broad_classes_and_go(all_conjugates_overlap_df)
  
  validation_candidates <- filter_candidates_expression(
    hits_df = all_conjugates_overlap_classified,
    expression_threshold = 0,
    method = "AND"
    )$kept
  
  # validation_hits <- simple_validation_selection(validation_candidates, n = 290)
  validation_hits <- select_validation_hits_balanced_classes(validation_candidates, n = 290)$selected_df
  validation_hits <- calculate_category_based_Z_score(validation_hits)
  
  saveRDS(all_conjugates_overlap_classified, file.path(dual_rep_dir, "all_conjugates_overlap_classified.rds"))
  saveRDS(validation_hits, file.path(dual_rep_dir, "validation_hits.rds"))
} else {
  all_conjugates_overlap_classified <- readRDS(file.path(dual_rep_dir, "all_conjugates_overlap_classified.rds"))
  validation_hits <- readRDS(file.path(dual_rep_dir, "validation_hits.rds"))
}
```

``` {r compare_hits_against_cardio_lib, eval=False}
cardio_library <- readRDS(file.path("/g/steinmetz/link/Gene_list_finalize","cardio_sgRNA_df_final.rds")) %>%
  distinct(entrez_id, .keep_all = TRUE) %>%
  select(entrez_id, gene)

# unique Entrez IDs from cardio_library
cardio_entrez <- unique(cardio_library$entrez_id)

genes_not_in_cardio <- all_conjugates_overlap_classified %>%
  filter(!(entrez %in% cardio_entrez))

# summary for plotting
plot_df <- all_conjugates_overlap_classified %>%
  mutate(in_cardio = entrez %in% cardio_entrez) %>%
  group_by(category) %>%
  summarise(
    n_entries = n(),
    n_in_cardio = sum(in_cardio, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(n_entries, n_in_cardio),
    names_to = "bar_type",
    values_to = "count"
  ) %>%
  mutate(
    bar_type = recode(
      bar_type,
      n_entries = "All entries",
      n_in_cardio = "In cardio_library"
    )
  )

# barplot
ggplot(plot_df, aes(x = category, y = count, fill = bar_type)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(
    aes(label = count),
    position = position_dodge(width = 0.8),
    vjust = -0.3,
    size = 3
  ) +
  labs(
    x = "Category",
    y = "Number of entries",
    fill = NULL
  ) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

write_xlsx(genes_not_in_cardio, file.path(dual_rep_dir, "Hits_not_in_cardio_specific_lib.xlsx"))

```

``` {r histogram_Z-score_distribution, eval=False}
plot_significanceZ_histogram(all_conjugates_overlap_df, normalize = TRUE, plot_title = "Explorative Hits", fill_color = "forestgreen")
plot_significanceZ_histogram(validation_hits, normalize = TRUE, plot_title = "Class Distribution Validation selection", fill_color = "steelblue")
plot_significanceZ_histogram(all_conjugates_overlap_df, normalize = FALSE, plot_title = "Explorative Hits", fill_color = "forestgreen")
plot_significanceZ_histogram(validation_hits, normalize = FALSE, plot_title = "Class Distribution Validation selection", fill_color = "steelblue")
```

``` {r horizontal_barplot_broad_class_percentage, eval=False}
# All Hits
plot_class_distribution(
  df_all = all_conjugates_overlap_classified,
  df_validation = validation_hits
)
# All Hits
plot_class_distribution(
  df_all = all_conjugates_overlap_classified,
  df_validation = validation_hits,
  sign_split = TRUE
)
# shared Hits
plot_class_distribution(
  df_all = all_conjugates_overlap_classified %>% filter(!category %in% c("unique_PA", "unique_DCA", "unique_GalNAc")),
  df_validation = validation_hits %>% filter(!category %in% c("unique_PA", "unique_DCA", "unique_GalNAc")),
  sign_split = TRUE
)
# Unique Hits
plot_class_distribution(
  df_all = all_conjugates_overlap_classified %>% filter(category %in% c("unique_PA", "unique_DCA", "unique_GalNAc")),
  df_validation = validation_hits %>% filter(category %in% c("unique_PA", "unique_DCA", "unique_GalNAc")),
  sign_split = TRUE
)

```

``` {r filter_candidates_functions, eval=False}


validation_hits_s1 <- filter_candidates_expression(
  hits_df = explorative_hits,
  expression_threshold = 0,
  method = "AND"
  )$kept
hits_removed_bc_expression <- filter_candidates_expression(
  hits_df = explorative_hits,
  expression_threshold = 0,
  method = "AND"
  )$removed

if (first_time){
  pathways_Z_enriched <- perform_pathway_analysis(
    file_info_suffix,
    FDR_threshold = 0.05,
    go_enrichment = "ALL",
    go_terms_qvalueCutoff = 0.05)
  pathways_Z_enriched_all <- pathways_Z_enriched$gsea_all
  pathways_Z_enriched_consensus <- pathways_Z_enriched$gsea_consensus
  
  
  pathways_hit_enriched <- perform_pathway_analysis(
    file_info_suffix,
    FDR_threshold = 0.05,
    go_enrichment = "ALL",
    method = "enrichGO",
    consensus = TRUE,
    enrichGO_consensus_df = explorative_hits,
    go_terms_qvalueCutoff = 0.05)
  
  
  ####
  # Manually adding GO terms, that sound intersting, but did not pass consensus calling
  manual_pathway_additions <- c(
    "protein localization to vacuole",
    "endocytic recycling",
    "lysosome localization",
    "regulation of ER to Golgi vesicle-mediated transport",
    "vacuolar localization",
    "vesicle transport along microtubule",
    "exocytic process"
  )
  ####
  
  pathways_combined <- bind_rows(
    pathways_hit_enriched,
    pathways_Z_enriched_consensus,
    pathways_Z_enriched_all %>% 
      filter(Description %in% manual_pathway_additions)) %>%
    distinct(ID, .keep_all = TRUE)
  
  gene_centrality <- get_gene_centrality(file_info_suffix, pathways_combined, use_local = TRUE) %>% 
    filter(!entrez %in% hits_removed_bc_expression$entrez)
  
  validation_results <- select_validation_genes(
    core_genes = gene_centrality,
    hits_df = explorative_hits,
    consensus_pathways = pathways_combined,
    top_n = 1
  )
  
  val_validation_genes <- validation_results$validation_genes
  val_hits_no_pathway <- validation_results$hits_without_pathway
  val_not_picked <- validation_results$hits_not_selected
  val_pathways_no_hits <- validation_results$pathways_without_hits
  
  pathway_analysis_results <- list(
    val_validation_genes = val_validation_genes,
    val_hits_no_pathway = val_hits_no_pathway,
    val_not_picked = val_not_picked,
    val_pathways_no_hits = val_pathways_no_hits,
    gene_centrality = gene_centrality,
    pathways_combined = pathways_combined,
    pathways_Z_enriched_all = pathways_Z_enriched_all,
    pathways_Z_enriched_consensus = pathways_Z_enriched_consensus
  )
  
  saveRDS(
    pathway_analysis_results,
    file.path(rds_output_folder,
              paste0(file_info_suffix,"_pathway_analysis_results.rds")))
} else {
  pathway_analysis_results <- readRDS(file.path(
    rds_output_folder,
    paste0(file_info_suffix,"_pathway_analysis_results.rds")))

  val_validation_genes <- pathway_analysis_results$val_validation_genes
  val_hits_no_pathway <- pathway_analysis_results$val_hits_no_pathway
  val_not_picked <- pathway_analysis_results$val_not_picked
  val_pathways_no_hits <- pathway_analysis_results$val_pathways_no_hits
  gene_centrality <- pathway_analysis_results$gene_centrality
  pathways_combined <- pathway_analysis_results$pathways_combined
  pathways_Z_enriched_all <- pathway_analysis_results$pathways_Z_enriched_all
  pathways_Z_enriched_consensus <- pathway_analysis_results$pathways_Z_enriched_consensus
}

# saveRDS(pathway_analysis_results_PA, "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/pathway_analysis_results_PA")
# saveRDS(pathway_analysis_results_DCA, "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/pathway_analysis_results_DCA")
# saveRDS(pathway_analysis_results_GALNAC, "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/pathway_analysis_results_GALNAC")

val_validation_genes_combined <- bind_rows(
  pathway_analysis_results_PA$val_validation_genes %>%
    mutate(pathway_analysis = "PA"),
  pathway_analysis_results_DCA$val_validation_genes %>%
    mutate(pathway_analysis = "DCA"),
  pathway_analysis_results_GALNAC$val_validation_genes %>%
    mutate(pathway_analysis = "GALNAC")) %>% 
  distinct(entrez, .keep_all = TRUE)

val_not_picked_combined <- bind_rows(
  pathway_analysis_results_PA$val_not_picked %>%
    mutate(pathway_analysis = "PA"),
  pathway_analysis_results_DCA$val_not_picked %>%
    mutate(pathway_analysis = "DCA"),
  pathway_analysis_results_GALNAC$val_not_picked%>%
    mutate(pathway_analysis = "GALNAC")) %>%
  distinct(entrez, .keep_all = TRUE) %>% 
  filter(!entrez %in% val_validation_genes_combined$entrez)

val_hits_no_pathway_combined <- bind_rows(
  pathway_analysis_results_PA$val_hits_no_pathway %>%
    mutate(pathway_analysis = "PA"),
  pathway_analysis_results_DCA$val_hits_no_pathway %>%
    mutate(pathway_analysis = "DCA"),
  pathway_analysis_results_GALNAC$val_hits_no_pathway%>%
    mutate(pathway_analysis = "GALNAC")) %>%
  distinct(entrez, .keep_all = TRUE) %>% 
  filter(!entrez %in% val_validation_genes_combined$entrez) %>% 
  filter(!entrez %in% val_not_picked_combined$entrez)


category_summary_df <- full_join(
  all_conjugates_overlap_df %>% count(category, name = "n_all"),
  all_conjugates_overlap_df %>% 
    filter(entrez %in% validation_hits$entrez) %>%
    count(category, name = "n_validation"),
  by = "category"
) %>%
  full_join(
    all_conjugates_overlap_df %>% 
      filter(entrez %in% val_hits_no_pathway_combined$entrez) %>% 
      count(category, name = "n_no_pathway"),
    by = "category"
  ) %>%
  full_join(
    all_conjugates_overlap_df %>% 
      filter(entrez %in% val_not_picked_combined$entrez) %>% 
      count(category, name = "n_not_picked"),
    by = "category"
  ) %>%
  mutate(across(-category, ~replace_na(., 0)))

plot_df <- category_summary_df %>%
  tidyr::pivot_longer(
    cols = c(n_all, n_validation, n_no_pathway, n_not_picked),
    names_to = "group",
    values_to = "count"
  ) %>%
  mutate(
    group = factor(
      group,
      levels = c("n_all", "n_validation", "n_no_pathway", "n_not_picked"),
      labels = c("All", "Validation", "No pathway", "Not picked")
    )
  )

p <- ggplot(plot_df, aes(x = group, y = count, fill = group)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = count), vjust = -0.3, size = 3.5) +
  facet_grid(category ~ ., scales = "free_y", switch = "y") +
  scale_fill_manual(values = c(
    "All" = "grey40",
    "Validation" = "forestgreen",
    "No pathway" = "tomato",
    "Not picked" = "steelblue"
  )) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
  labs(
    x = NULL,
    y = "Count"
  ) +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.placement = "outside",
    strip.background = element_blank(),
    panel.spacing.y = unit(1.2, "lines")
  )

ggsave(
  "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/category_barplots.png",
  plot = p,
  width = 7,
  height = 12
)


draw_category_venn <- function(df, value_col, main_title = value_col) {
  d <- setNames(df[[value_col]], df$category)
  
  venn.plot <- draw.triple.venn(
    area1 = d["unique_DCA"] + d["shared_DCA_PA"] + d["shared_DCA_GalNAc"] + d["shared_all_three"],
    area2 = d["unique_PA"] + d["shared_DCA_PA"] + d["shared_PA_GalNAc"] + d["shared_all_three"],
    area3 = d["unique_GalNAc"] + d["shared_DCA_GalNAc"] + d["shared_PA_GalNAc"] + d["shared_all_three"],
    
    n12 = d["shared_DCA_PA"] + d["shared_all_three"],
    n13 = d["shared_DCA_GalNAc"] + d["shared_all_three"],
    n23 = d["shared_PA_GalNAc"] + d["shared_all_three"],
    n123 = d["shared_all_three"],
    
    category = c("DCA", "PA", "GalNAc"),
    fill = c("tomato", "steelblue", "forestgreen"),
    alpha = 0.5,
    cex = 1.3,
    cat.cex = 1.2,
    cat.pos = c(-20, 20, 180),
    main = main_title
  )
  
  grid.newpage()
  grid.draw(venn.plot)
}
draw_category_venn(category_summary_df, "n_all", "All")
draw_category_venn(category_summary_df, "n_all", "All")


draw_category_venn(category_summary_df, "n_all", "All")
draw_category_venn(category_summary_df, "n_validation", "Validation")
draw_category_venn(category_summary_df, "n_no_pathway", "No pathway")
draw_category_venn(category_summary_df, "n_not_picked", "Not picked")
```

``` {r filter_candidates, eval=False}
nsZh <- TRUE

val_validation_genes_combined <- val_validation_genes_combined %>%
  left_join(
    all_conjugates_overlap_df %>% select(entrez, significanceZ),
    by = "entrez",
    suffix = c("", ".new")
  ) %>%
  mutate(significanceZ = significanceZ.new) %>%
  select(-significanceZ.new)




plot_significanceZ_histogram(val_validation_genes_combined, normalize = nsZh, plot_title = "GO Term Hits", fill_color = "darkgreen")
plot_significanceZ_histogram(val_not_picked, normalize = nsZh, plot_title = "Hits not chosen", fill_color = "orange")
plot_significanceZ_histogram(val_hits_no_pathway, normalize = nsZh, plot_title = "Hits that don't match enriched GO Terms", fill_color = "red")

```



``` {r Comparing_GO-terms_of_different_moieties, eval=FALSE}
source(file.path(project_root_dir, "functions", "zzz_source_all.R"))

dual_rep_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/"
all_conjugates_overlap_classified <- readRDS(file.path(dual_rep_dir, "all_conjugates_overlap_classified.rds"))


# pathways_hit_enriched_by_cat <- run_pathway_by_category(
#   df = all_conjugates_overlap_classified,
#   split_pos_neg = TRUE,
#   include_shared_hits = c("shared_all_three", "DCA", "PA", "GalNAc"),
#   plot_enrichment = TRUE,
#   image_save_folder = file.path(dual_rep_dir, "GO_enrichment_plots/CC-pos_neg_split-shared_included"),
#   go_enrichment = "CC",
#   file_info_suffix = file_info_suffix,
#   method = "enrichGO"
# )
if (first_time){
  #careful this code takes a while to run ~1h
  all_pathway_results <- run_all_pathway_combinations( 
    df = all_conjugates_overlap_classified,
    dual_rep_dir = dual_rep_dir,
    file_info_suffix = file_info_suffix,
    use_prior_results = TRUE,
  )
  saveRDS(all_pathway_results, file.path(dual_rep_dir, "GO_enrichment_plots/all_go_results.rds"))
} else {
  all_pathway_results <- readRDS(file.path(dual_rep_dir, "GO_enrichment_plots/all_go_results.rds"))
}

check_go_results_have_png <- function(root_dir) {
  
  go_result_files <- list.files(
    path = root_dir,
    pattern = "_go_results\\.rds$",
    recursive = TRUE,
    full. <- s = TRUE
  )
  
  if (length(go_result_files) == 0) {
    stop("No '*_go_results.rds' files found.")
  }
  
  has_rows <- function(x) {
    if (is.null(x)) {
      return(FALSE)
    }
    
    if (is.data.frame(x)) {
      return(nrow(x) > 0)
    }
    
    # fallback for objects that can be turned into data.frame
    out <- tryCatch(as.data.frame(x), error = function(e) NULL)
    if (!is.null(out)) {
      return(nrow(out) > 0)
    }
    
    FALSE
  }
  
  check_df <- lapply(go_result_files, function(f) {
    obj <- readRDS(f)
    
    non_empty <- has_rows(obj)
    
    expected_png <- file.path(
      dirname(f),
      sub("_go_results\\.rds$", ".png", basename(f))
    )
    
    png_exists <- file.exists(expected_png)
    
    data.frame(
      go_results_file = f,
      expected_png = expected_png,
      non_empty = non_empty,
      png_exists = png_exists,
      check_pass = (!non_empty) | png_exists,
      stringsAsFactors = FALSE
    )
  }) %>%
    dplyr::bind_rows()
  
  return(check_df)
}
png_check <- check_go_results_have_png(dual_rep_dir)
png_missing <- png_check %>% filter(check_pass == "FALSE")

run_pathway_by_category(all_conjugates_overlap_classified,
                                    file_info_suffix,
                                    category_col = "category",
                                    FDR_threshold = 0.05,
                                    go_enrichment = "CC",
                                    method = "enrichGO",
                                    consensus = TRUE,
                                    go_terms_qvalueCutoff = 0.05,
                                    include_shared_hits = NULL,
                                    split_pos_neg = TRUE,
                                    plot_enrichment = TRUE,
                                    image_save_folder = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/GO_enrichment_plots/CC-pos_neg_split-no_shared")
	
run_pathway_by_category(all_conjugates_overlap_classified,
                                    file_info_suffix,
                                    category_col = "category",
                                    FDR_threshold = 0.05,
                                    go_enrichment = "MF",
                                    method = "enrichGO",
                                    consensus = TRUE,
                                    go_terms_qvalueCutoff = 0.05,
                                    include_shared_hits = NULL,
                                    split_pos_neg = FALSE,
                                    plot_enrichment = TRUE,
                                    image_save_folder = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/GO_enrichment_plots/MF-no_pos_neg_split-no_shared")
run_pathway_by_category(all_conjugates_overlap_classified,
                                    file_info_suffix,
                                    category_col = "category",
                                    FDR_threshold = 0.05,
                                    go_enrichment = "MF",
                                    method = "enrichGO",
                                    consensus = TRUE,
                                    go_terms_qvalueCutoff = 0.05,
                                    include_shared_hits = NULL,
                                    split_pos_neg = TRUE,
                                    plot_enrichment = TRUE,
                                    image_save_folder = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/GO_enrichment_plots/MF-pos_neg_split-no_shared")
run_pathway_by_category(all_conjugates_overlap_classified,
                                    file_info_suffix,
                                    category_col = "category",
                                    FDR_threshold = 0.05,
                                    go_enrichment = "MF",
                                    method = "enrichGO",
                                    consensus = TRUE,
                                    go_terms_qvalueCutoff = 0.05,
                                    include_shared_hits = TRUE,
                                    split_pos_neg = TRUE,
                                    plot_enrichment = TRUE,
                                    image_save_folder = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/GO_enrichment_plots/MF-pos_neg_split-shared_included")
diagnose_go_plot_missing <- function(go_results_file) {
  
  hits_file <- sub("_go_results\\.rds$", "_hits.rds", go_results_file)
  
  res_obj <- readRDS(go_results_file)
  hits_df <- if (file.exists(hits_file)) readRDS(hits_file) else NULL
  
  res_df <- tryCatch(as.data.frame(res_obj), error = function(e) NULL)
  
  out <- list(
    go_results_file = go_results_file,
    hits_file = hits_file,
    hits_file_exists = file.exists(hits_file),
    res_is_null = is.null(res_obj),
    res_df_is_null = is.null(res_df),
    res_nrow = if (!is.null(res_df)) nrow(res_df) else NA_integer_,
    has_required_cols = if (!is.null(res_df)) all(c("ID", "Description", "p.adjust") %in% colnames(res_df)) else FALSE,
    hits_is_null = is.null(hits_df),
    hits_nrow = if (!is.null(hits_df)) nrow(hits_df) else NA_integer_,
    has_GO_terms_col = if (!is.null(hits_df)) "GO_terms" %in% colnames(hits_df) else FALSE
  )
  
  if (!is.null(hits_df) && "GO_terms" %in% colnames(hits_df) && !is.null(res_df) && nrow(res_df) > 0) {
    go_lists <- strsplit(
      ifelse(is.na(hits_df$GO_terms), "", as.character(hits_df$GO_terms)),
      ";",
      fixed = TRUE
    )
    
    counts <- vapply(
      res_df$ID,
      function(go_id) {
        sum(vapply(go_lists, function(x) go_id %in% trimws(x), logical(1)))
      },
      numeric(1)
    )
    
    out$n_hits_with_nonmissing_GO_terms <- sum(!is.na(hits_df$GO_terms) & hits_df$GO_terms != "")
    out$n_result_terms = nrow(res_df)
    out$n_result_terms_with_Count_gt_0 = sum(counts > 0)
    out$max_Count = if (length(counts) > 0) max(counts) else NA_real_
    out$all_Count_zero = all(counts == 0)
    
    if (sum(counts > 0) > 0) {
      out$first_matching_terms <- head(res_df$ID[counts > 0], 10)
    } else {
      out$first_matching_terms <- character(0)
    }
  }
  
  return(out)
}
lapply(png_missing$go_results_file, diagnose_go_plot_missing)
``` 

``` {r make_STRINGI_images, eval=FALSE}
source(file.path(project_root_dir, "functions", "zzz_source_all.R"))
dual_rep_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/"

plot_string_from_saved_hits(dual_rep_dir) #careful this code takes a while to run ~1h


``` 

``` {r Hit_significance_heatmap, eval=FALSE}


dual_rep_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/"

heatmap_res <- plot_three_mean_hit_heatmap(
  PA_means = PA_means,
  DCA_means = DCA_means,
  GALNAC_means = GALNAC_means,
  all_conjugates_overlap_classified = all_conjugates_overlap_classified,
  value_col = "significanceZ",
  show_significance_stars = TRUE,
  show_clustering_tree = TRUE,
  show_colnames = TRUE,
  cluster_rows = TRUE,
  n_clusters = 14,
  create_png_path = file.path(dual_rep_dir, "full_hit_heatmap.png")
)
``` 

``` {r Hit_significance_UMAP, eval=FALSE}

dual_rep_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/"
umap_res <- plot_three_screen_significance_umap(
  PA_means = PA_means,
  DCA_means = DCA_means,
  GALNAC_means = GALNAC_means,
  all_conjugates_overlap_classified = all_conjugates_overlap_classified,
  color_by = "category",
  label_points = FALSE,
  create_png_path = file.path(dual_rep_dir, "exlpore_hit_UMAP.png"),
  use_only_overlap_genes = TRUE,
)
```

``` {r control_distribution_heatmap, eval=FALSE}
library(pheatmap)
library(grid)
# 1) Add RPM per (condition, sublib, sample)
count_df_long2 <- count_df_long %>%
  group_by(condition, sublib, sample) %>%
  mutate(RPM = (count / sum(count)) * 1e6) %>%
  ungroup()

# 2) Remove targeting guides (not relevant for heatmap)
count_df_long2 <- count_df_long2 %>%
  filter(group_category != "targeting")

# 3) Make a column ID for the heatmap columns
count_df_long2 <- count_df_long2 %>%
  mutate(col_id = paste(condition, sublib, sample, sep = "_"))

# 4) Ensure 1 RPM value per (sgRNA, col_id)
#    (If there are duplicates, sum them — or you could use mean)
heatmap_df <- count_df_long2 %>%
  group_by(sgRNA, col_id) %>%
  summarise(RPM = sum(RPM), .groups = "drop")

# 5) Pivot to wide format for heatmap input
heatmap_wide <- heatmap_df %>%
  pivot_wider(names_from = col_id, values_from = RPM, values_fill = 0) %>%
  arrange(sgRNA)

# 6) Convert to matrix with sgRNA as rownames
heatmap_mat <- as.matrix(heatmap_wide[, -1])
rownames(heatmap_mat) <- heatmap_wide$sgRNA

chunk_size <- 60

# (optional) order rows by variance so "interesting" ones appear first
row_var <- apply(heatmap_mat, 1, var)
heatmap_mat <- heatmap_mat[order(row_var, decreasing = TRUE), , drop = FALSE]

# split row indices into chunks of 50
idx_list <- split(seq_len(nrow(heatmap_mat)),
                  ceiling(seq_len(nrow(heatmap_mat)) / chunk_size))

# list of matrices, each with up to 50 rows
heatmap_chunks <- lapply(idx_list, function(idx) heatmap_mat[idx, , drop = FALSE])

for (k in seq_along(heatmap_chunks)) {
  row_start <- min(idx_list[[k]])
  row_end   <- max(idx_list[[k]])

  pheatmap::pheatmap(
    heatmap_chunks[[k]],
    scale = "none",
    fontsize_row = 6,
    main = paste0("Heatmap ", k, " (rows ", row_start, "–", row_end, ")")
  )
}

pdf_file <- get_file_path(results_output_folder, "heatmaps_chunks_60rows.pdf")

grDevices::pdf(pdf_file, width = 10, height = 18, onefile = TRUE)
pdf_dev <- grDevices::dev.cur()   # <- store the pdf device id

for (k in seq_along(heatmap_chunks)) {

  row_start <- min(idx_list[[k]])
  row_end   <- max(idx_list[[k]])

  ph <- pheatmap::pheatmap(
    heatmap_chunks[[k]],
    scale = "none",
    main = paste0("Heatmap ", k, " (rows ", row_start, "–", row_end, ")"),
    fontsize_row = 6,
    silent = TRUE
  )

  grDevices::dev.set(pdf_dev)     # <- force drawing into the PDF device
  grid::grid.newpage()
  grid::grid.draw(ph$gtable)

  invisible(ph)                  # <- prevents knitr from printing it
}

grDevices::dev.off()


```


```{r input_vs_plasmid, eval=FALSE}
res <- scatter_merged_vs_input(merged_sgRNA_df, count_df_long,
                               normalization = "CPM",
                               group_category = "all",
                               point_size = 0.5)
res <- scatter_merged_vs_input(merged_sgRNA_df, count_df_long,
                               normalization = "CPM",
                               group_category = "all",
                               remove_essential = TRUE,
                               point_size = 0.5)

res <- scatter_merged_vs_input(merged_sgRNA_df, count_df_long,
                               normalization = "CPM",
                               group_category = c("targeting"),
                               sublib_chosen = "sublib_4",
                               point_size = 0.5)
res <- scatter_merged_vs_input(merged_sgRNA_df, count_df_long,
                               normalization = "CPM",
                               group_category = c("non_targeting_control"),
                               sublib_chosen = "sublib_4",
                               mark_outlier = FALSE,
                               point_size = 0.5)
res <- scatter_merged_vs_input(merged_sgRNA_df, count_df_long,
                               normalization = "none",
                               group_category = c("non_targeting_control"),
                               sublib_chosen = "sublib_4",
                               mark_outlier = FALSE,
                               point_size = 0.5)

```

``` {r Manhattan_plot, eval=False}
Hits_current_settings <- add_info_wrapper(file_info_suffix)
if (length(include_controls_list) > 0) {
  for (control_gene in include_controls_list) {
    # here we remove everything after the last _, so stuff like AAVS1_9 and
    # AAVS1_13 are both treated as AAVS1
    control_gene <- sub("_[^_]*$", "", control_gene)
    Hits_current_settings$symbol[grepl(control_gene, Hits_current_settings$entrez)] <- control_gene
  }
  # Filter out rows where the 'symbol' is in the include_controls_list
  Hits_current_settings_no_controls <- Hits_current_settings %>% 
    filter(!symbol %in% include_controls_list | is.na(symbol))
}
Hits_df <- Hits_current_settings %>% 
  filter(significanceZ <= 0) %>% 
  mutate(logP = -log10(FDR))

# Define the cluster and protein names
clusters <- data.frame(
  cluster_name = c(
    "Protein ufmylation", 
    "mTORC1-mediated signalling (Ragulator complex/RagC protein)", 
    "vesicles thetering complex (HOPS complex)", 
    "ER membrane insertion complex, and GET complex", 
    "GTPase activating proteins (GAP)", 
    "Regulation of lysosome size (BLOC-1 and BORC complexes)", 
    "Other VPS proteins", 
    "Other Rab proteins"
  ),
  color = c(
    "#ff0000", "#ffe065", "#47b200", "#65ffa3", 
    "#008eb2", "#6566ff", "black", "black"
  ),
  proteins = c(
    "UFL1, CDK5RAP3, UFM1, UBA5", 
    "LAMTOR1, RRAGC, LAMTOR5", 
    "VPS39, VPS33A, VPS41", 
    "BAG6, GET1, GET4", 
    "RAB3GAP1, RAB3GAP2, TBC1D20", 
    "BLOC1S1, SNAPIN", 
    "VPS13C, VPS53", 
    "RAB18"
  )
)

# Flatten protein names into a list
protein_list <- unlist(strsplit(clusters$proteins, ", "))

# Create a new column in Hits_df to mark the cluster for each protein
Hits_df$cluster <- sapply(Hits_df$symbol, function(x) {
  # Check which cluster the protein belongs to
  match <- which(sapply(clusters$proteins, function(proteins) x %in% strsplit(proteins, ", ")[[1]]))
  if (length(match) > 0) {
    return(clusters$cluster_name[match])
  } else {
    return("Other")  # Default category for proteins not listed
  }
})

Hits_df$cluster_color <- sapply(Hits_df$cluster, function(x) {
  # Check if the cluster is in the list, else assign grey
  if (x %in% clusters$cluster_name) {
    return(clusters$color[clusters$cluster_name == x])
  } else {
    return("grey")  # Assign grey to other genes
  }
})
# Define the clusters excluding "Other"
clusters_excluding_other <- unique(Hits_df$cluster[Hits_df$cluster != "Other"])

# Define the number of clusters excluding "Other"
num_clusters <- length(clusters_excluding_other)

# Define the size of each cluster's range (divide 1000 by the number of clusters - 1)
range_size <- 1000 / (num_clusters)

# Initialize a vector to store the ranks
rank_vector <- numeric(nrow(Hits_df))

# Loop through each cluster excluding "Other"
for (cluster_name in clusters_excluding_other) {
  # Get the rows for the current cluster
  cluster_rows <- Hits_df$cluster == cluster_name
  
  # Calculate the range for this cluster
  cluster_start <- (which(clusters_excluding_other == cluster_name) - 1) * range_size + 1
  cluster_end <- cluster_start + range_size - 1
  
  # Assign random ranks within this cluster's range
  rank_vector[cluster_rows] <- sample(cluster_start:cluster_end, sum(cluster_rows), replace = TRUE)
}

# Assign ranks to "Other" cluster (random ranks from 1 to 1000)
rank_vector[Hits_df$cluster == "Other"] <- sample(1:1000, sum(Hits_df$cluster == "Other"), replace = TRUE)

# Add the new rank column to the Hits_df dataframe
Hits_df$rank <- rank_vector

# Create the Manhattan plot
p <- ggplot(Hits_df, aes(x = rank, y = significanceZ, color = cluster_color)) +
  geom_point(aes(size = ifelse(cluster != "Other", 3, 1)), alpha = 0.8) +
  scale_size_continuous(range = c(1, 5)) +
  scale_y_reverse() +
  #scale_y_log10() +
  scale_color_identity() +  # Use the actual colors defined for each cluster
  labs(
    title = "Non-Essential Gene Groups",
    y = "significanceZ"
  ) +
  theme_classic() +
  theme(
    legend.position = "none",  # Remove legend for clusters (you can add one if needed)
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    
    # Remove X-axis and its label, ticks, and lines
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank()
  ) + 
  # Use geom_text_repel to avoid label overlap and label non-"Other" points
  geom_text_repel(
    data = subset(Hits_df, cluster != "Other"),  # Only label non-"Other" points
    aes(x = rank, y = significanceZ, label = symbol),
    color = "black",
    size = 3,
    box.padding = 0.1,  # Increase to move text away from points if needed
    point.padding = 0.2,  # Adjust text padding around points
    max.overlaps = 10,   # Adjust this value to limit the number of overlaps
    segment.color = NA,  # No lines connecting text and point
    vjust = -2          # Position text above the dot
  )

# Display the plot
print(p)

cluster_labels <- clusters
# Create a blank plot with adjusted limits
# Create a blank plot with adjusted limits
# Create a blank plot with adjusted limits
plot(
  x = 1, y = 1, type = "n", 
  xlim = c(0, 500), ylim = c(1, nrow(cluster_labels) + 1),  # Increase xlim to make space for the text
  xlab = "", ylab = "", 
  axes = FALSE,  # Remove axes
  main = "Cluster Names and Colors"
)

# Plot each cluster with a single dot and its name next to it
for (i in 1:nrow(cluster_labels)) {
  # Plot the dot on the left (x = 0) for each cluster (ensure the dot is visible)
  points(1, nrow(cluster_labels) - i + 1, pch = 16, cex = 2, col = cluster_labels$color[i])  
  # Add the cluster name to the right of the dot (slightly shifted right)
  text(1.5, nrow(cluster_labels) - i + 1, cluster_labels$cluster_name[i], pos = 4, cex = 1.2)  
}

```

``` {r test_1s, eval=FALSE}
# 1. Count the number of 1s in the input, lower, and upper columns individually
count_1s_input <- sum(maude_counts_df$input == 1)
count_1s_lower <- sum(maude_counts_df$lower == 1)
count_1s_upper <- sum(maude_counts_df$upper == 1)

# Print individual counts
cat("Count of 1's in 'input':", count_1s_input, "\n")
cat("Count of 1's in 'lower':", count_1s_lower, "\n")
cat("Count of 1's in 'upper':", count_1s_upper, "\n")

# 2. Count rows where all three columns are 1 (input, lower, upper)
count_all_ones <- sum(maude_counts_df$input == 1 & maude_counts_df$lower == 1 & maude_counts_df$upper == 1)
cat("Rows where all three columns are 1:", count_all_ones, "\n")

# 3. Separate counts based on 'isNontargeting' column (TRUE and FALSE)

# For rows where 'isNontargeting' is TRUE
df_true <- maude_counts_df[maude_counts_df$isNontargeting == TRUE, ]
count_1s_input_true <- sum(df_true$input == 1)
count_1s_lower_true <- sum(df_true$lower == 1)
count_1s_upper_true <- sum(df_true$upper == 1)
count_all_ones_true <- sum(df_true$input == 1 & df_true$lower == 1 & df_true$upper == 1)

cat("\nFor isNontargeting == TRUE:\n")
cat("Count of 1's in 'input':", count_1s_input_true, "\n")
cat("Count of 1's in 'lower':", count_1s_lower_true, "\n")
cat("Count of 1's in 'upper':", count_1s_upper_true, "\n")
cat("Rows where all three columns are 1:", count_all_ones_true, "\n")

# For rows where 'isNontargeting' is FALSE
df_false <- maude_counts_df[maude_counts_df$isNontargeting == FALSE, ]
count_1s_input_false <- sum(df_false$input == 1)
count_1s_lower_false <- sum(df_false$lower == 1)
count_1s_upper_false <- sum(df_false$upper == 1)
count_all_ones_false <- sum(df_false$input == 1 & df_false$lower == 1 & df_false$upper == 1)

cat("\nFor isNontargeting == FALSE:\n")
cat("Count of 1's in 'input':", count_1s_input_false, "\n")
cat("Count of 1's in 'lower':", count_1s_lower_false, "\n")
cat("Count of 1's in 'upper':", count_1s_upper_false, "\n")
cat("Rows where all three columns are 1:", count_all_ones_false, "\n")

```


``` {r Roche_compare, eval=FALSE}

if (length(include_controls_list) > 0) {
  for (control_gene in include_controls_list) {
    # here we remove everything after the last _, so stuff like AAVS1_9 and
    # AAVS1_13 are both treated as AAVS1
    control_gene <- sub("_[^_]*$", "", control_gene)
    Hits_current_settings$symbol[grepl(control_gene, Hits_current_settings$entrez)] <- control_gene
  }
  # Filter out rows where the 'symbol' is in the include_controls_list
  Hits_current_settings_no_controls <- Hits_current_settings %>% 
    filter(!symbol %in% include_controls_list | is.na(symbol))
}


conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::rename)
conflicts_prefer(dplyr::filter)
dual_rep_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/"

all_conjugates_overlap_classified <- readRDS(file.path(dual_rep_dir, "all_conjugates_overlap_classified.rds"))

roche_folder <- make_clean_dir(dual_rep_dir, "/Roche_compare")

sup_3_path <- get_file_path(roche_folder, "41467_2025_61039_MOESM5_ESM-3.xls")
sup_3 <- read_excel(sup_3_path)

sup_3$GeneID <- gsub("GeneID:", "", sup_3$GeneID)
sup_3_rename <- sup_3 %>%
  rename(
    entrez = GeneID,
    Rho_enriched = Rho_Bulk_to_Reference_enriched,
    MedianLFC_enriched = MedianLFC_Bulk_to_Reference_enriched,
    AdjPval_enriched = AdjPval_Bulk_to_Reference_enriched,
    Pval_enriched = Pval_Bulk_to_Reference_enriched,
    RhoRank_enriched = RhoRank_Bulk_to_Reference_enriched,
    Rho_depleted = Rho_Bulk_to_Reference_depleted,
    MedianLFC_depleted = MedianLFC_Bulk_to_Reference_depleted,
    AdjPval_depleted = AdjPval_Bulk_to_Reference_depleted,
    Pval_depleted = Pval_Bulk_to_Reference_depleted,
    RhoRank_depleted = RhoRank_Bulk_to_Reference_depleted
  ) %>% 
  select(
    entrez,
    Rho_enriched,
    MedianLFC_enriched,
    AdjPval_enriched,
    Pval_enriched,
    RhoRank_enriched,
    Rho_depleted,
    MedianLFC_depleted,
    AdjPval_depleted,
    Pval_depleted,
    RhoRank_depleted
  )
compare_df <- merge(sup_3_rename, merged_sgRNA_df, by = "entrez") %>% 
  dplyr::distinct(entrez, .keep_all = TRUE) %>% 
###############################################################################
# Venn Diagramm as requested by Nico
nico_compare_df <- sup_3 %>% 
  rename(Rho_pos = Rho_Round1Pos_to_Bulk_enriched,
         Rho_neg = Rho_Round1Neg_to_Bulk_enriched,
         entrez = GeneID,
         symbol = Symbol,
         LFC_pos = MedianLFC_Round1Pos_to_Bulk_enriched,
         LFC_neg = MedianLFC_Round1Neg_to_Bulk_enriched) %>% 
  select(entrez,
         Rho_pos, Rho_neg, LFC_pos,LFC_neg)
nico_compare_df <- merge(nico_compare_df, merged_sgRNA_df, by = "entrez") %>% 
  dplyr::distinct(entrez, .keep_all = TRUE) %>% 
  select(entrez, symbol,
         Rho_pos, Rho_neg, LFC_pos,LFC_neg)

  

# nico_hits_intersect_top_neg <- nico_compare_df %>% filter(significanceZ > 0 & FDR <= 0.05 & Rho_neg <= 0.0001 & 2^LFC_neg >= 1.5)
# nico_hits_intersect_low_pos <- nico_compare_df %>% filter(significanceZ < 0 & FDR <= 0.05 & Rho_pos <= 0.0001 & 2^LFC_pos >= 1.5)
# 
# our_hits_df <- nico_compare_df %>% filter(significanceZ > 0 & FDR <= 0.05)
# roche_hits_df <- nico_compare_df %>% filter(Rho_neg <= 0.0001 & 2^LFC_neg >= 1.5)

our_hits_df <- nico_compare_df 
roche_hits_df <- nico_compare_df %>%
  filter(
    (Rho_neg <= 0.0001 & 2^LFC_neg >= 1.5) |
    (Rho_pos <= 0.0001 & 2^LFC_pos >= 1.5)
  )

our_hits <- all_conjugates_overlap_classified$entrez
roche_hits <- roche_hits_df$entrez

overlap_our_hits_roche_hits <- merged_sgRNA_df %>% 
  filter(
    (entrez %in% our_hits) &
    (entrez %in% roche_hits)
  ) %>%
  dplyr::distinct(entrez, .keep_all = TRUE) %>% 
  select(entrez) %>% 
  mutate(entrez = as.character(entrez)) %>% 
  left_join(all_conjugates_overlap_classified, by = "entrez") %>% 
  select(entrez, symbol,
         significanceZ, significanceZ_PA,significanceZ_DCA,significanceZ_GalNAc,
         category, broad_class,
         essential_HepG2,essential_general,essential_liver,
         HepG2_CPM_Xie, HepG2_CPM_Battisti, HepG2_14B11_CPM_Battisti,
         GO_terms, GO_ontology, GO_terms_expanded) %>% 
  rename(sigZ_PA = significanceZ_PA,
         sigZ_DCA = significanceZ_DCA,
         sigZ_GAL = significanceZ_GalNAc)

write_xlsx(overlap_our_hits_roche_hits, "/g/steinmetz/link/sharing/20_Hits_overlap_between_roche_and_us.xlsx")

# Create the Venn as a grob (plot object)
venn_2 <- venn.diagram(
  x = list(
    our_hits = our_hits,
    roche_hits = roche_hits
  ),
  filename = NULL,          # NULL = return a grid object instead of saving to file
  fill = c("purple", "cyan"),
  alpha = 0.5,              # transparency
  cex = 1.5, 
  cat.pos  = c(0, 0),      # 90° = above each circle
  cat.dist = c(0.05, 0.05),  # distance from the circle border
  cat.cex  = 1.5# size of set names
)

# Draw it in the plotting window
grid.newpage()
grid.draw(venn_2)


# our_hits_df <- nico_compare_df %>% filter(significanceZ < 0 & FDR <= 0.05)
# roche_hits_df <- nico_compare_df %>% filter(Rho_pos <= 0.0001 & 2^LFC_pos >= 1.5)
# 
# our_hits <- our_hits_df$entrez
# roche_hits <- roche_hits_df$entrez
# 
# 
# # Create the Venn as a grob (plot object)
# venn_2 <- venn.diagram(
#   x = list(
#     our_hits = our_hits,
#     roche_hits = roche_hits
#   ),
#   filename = NULL,          # NULL = return a grid object instead of saving to file
#   fill = c("red", "yellow"),
#   alpha = 0.5,              # transparency
#   cex = 1.5, 
#   cat.pos  = c(0, 0),      # 90° = above each circle
#   cat.dist = c(0.05, 0.05),  # distance from the circle border
#   cat.cex  = 1.5# size of set names
# )
# 
# # Draw it in the plotting window
# grid.newpage()
# grid.draw(venn_2)
###############################################################################





waterfall_roche_compare(all_conjugates_overlap_classified, nico_compare_df)



################################################################################
upper_cut <- compare_df %>% 
  filter(significanceZ > 0)
under_cut <- compare_df %>% 
  filter(significanceZ < 0)

significant_df <- compare_df %>%
  filter(FDR <= 0.05 | AdjPval_enriched <= 0.05 | AdjPval_depleted <= 0.05)

significanter_df <- compare_df %>%
  filter((FDR <= 0.05 & AdjPval_enriched <= 0.05) | (AdjPval_depleted <= 0.05 & FDR <= 0.05))

colnames(compare_df)
ggplot(compare_df, aes(x = MedianLFC_enriched, y = significanceZ)) +
  geom_point() +  # Add points for scatter plot
  labs(x = "Median LFC Enriched", y = "Significance Z", 
       title = "compare_df: Median LFC Enriched vs. Significance Z") +
  theme_bw()  # Optional: for a clean, minimalistic theme
ggplot(significant_df, aes(x = MedianLFC_enriched, y = significanceZ)) +
  geom_point() +  # Add points for scatter plot
  labs(x = "Median LFC Enriched", y = "Significance Z", 
       title = "significant_df: Median LFC Enriched vs. Significance Z") +
  theme_bw()  # Optional: for a clean, minimalistic theme
# Calculate Spearman rank correlation between two columns
spearman_corr <- cor.test(compare_df$AdjPval_enriched, compare_df$FDR, method = "spearman")
# Print the result
print(spearman_corr)

spearman_corr <- cor.test(compare_df$Pval_enriched, compare_df$p.value, method = "spearman")
# Print the result
print(spearman_corr)

spearman_corr <- cor.test(compare_df$MedianLFC_enriched, compare_df$significanceZ, method = "spearman")
# Print the result
print(spearman_corr)
spearman_corr <- cor.test(significant_df$MedianLFC_depleted, significant_df$significanceZ, method = "spearman")
# Print the result
print(spearman_corr)

cor_df <- compare_df %>% select(-c(symbol, entrez, numGuides, seq, sgRNA, ))
cor_matrix <- cor(cor_df, method = "pearson")



make_cor_heatmap(compare_df, "compare_df")
make_cor_heatmap(under_cut, "under_cut")
make_cor_heatmap(upper_cut, "upper_cut")
print(make_cor_heatmap(significant_df, "significant_df"))

excel_file_path <- get_file_path(output_folder,"roche_us_overlap.xlsx")
write_xlsx(significanter_df, excel_file_path)
```

``` {r Autophagy_compare, eval=FALSE}
autophagy_path <- get_file_path(roche_folder, "Autophagy_Genes.xlsx")
autophagy_df <- read_excel(autophagy_path)
autophagy_compare_df <- merge(autophagy_df, Hits_current_settings, by = "entrez")

print(setdiff(autophagy_df$Autophagy_gene, autophagy_compare_df$Autophagy_gene))
```

``` {r sublibrary_compare, eval=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(ggplot2)
library(tibble)

norm_count_df_long <- normalize_count_df_long(count_df_long, norm_method = "control_median")

df <- norm_count_df_long %>%
  left_join(
    merged_sgRNA_df %>% select(sgrna_id, entrez),
    by = c("sgRNA" = "sgrna_id")
  ) %>%
  mutate(
    # EXACTLY what you asked: NA entrez -> sgRNA
    feature = if_else(is.na(entrez), sgRNA, as.character(entrez))

    # If you prefer matching across sublibs for the NA cases, use this instead:
    # feature = if_else(is.na(entrez), str_remove(sgRNA, "_\\d+$"), as.character(entrez))
  )
df_aggr <- df %>%
  group_by(condition, sublib, feature) %>%
  summarise(value = sum(count), .groups = "drop")
df_wide <- df_aggr %>%
  pivot_wider(names_from = sublib, values_from = value)

g <- df_wide %>% group_by(condition)
keys <- group_keys(g)$condition

cors_by_condition <- g %>%
  group_map(~{
    mat <- as.matrix(select(.x, where(is.numeric)))
    cor(mat, method = "spearman", use = "pairwise.complete.obs")
  }) %>%
  set_names(keys)



plot_cor_heatmap <- function(C, title = NULL) {
  as.data.frame(C) %>%
    rownames_to_column("lib1") %>%
    pivot_longer(-lib1, names_to = "lib2", values_to = "rho") %>%
    ggplot(aes(x = lib1, y = lib2, fill = rho)) +
    geom_tile() +
    geom_text(aes(label = sprintf("%.2f", rho)), size = 3) +
    coord_equal() +
    scale_fill_gradient2(limits = c(-1, 1)) +
    labs(title = title, x = NULL, y = NULL, fill = "Spearman\nρ") +
    theme_minimal(base_size = 12) +
    theme(
      panel.grid = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
}

# Create one plot per condition (cors_by_condition is a named list)
cor_plots <- imap(cors_by_condition, ~plot_cor_heatmap(.x, title = paste("Correlation heatmap:", .y)))
# 
# # Show one (e.g. input)
# cor_plots$input

# Or print all
walk(cor_plots, print)
```

``` {r Liangfu_Plots_1, eval=FALSE}
library(dplyr)
library(ggplot2)

count_df_long %>%
  filter(count > 0) %>%                      # log2 needs >0
  mutate(log2_count = log2(count)) %>%       # bins in log2 space
  ggplot(aes(x = log2_count)) +
  geom_histogram(bins = 50, boundary = 0) +
  facet_grid(condition ~ sublib) +
  labs(x = "log2(read count)", y = "Number of sgRNAs") +
  theme_bw()
# all unique (condition, sublib) combinations
groups <- count_df_long %>%
  distinct(condition, sublib)

for (i in seq_len(nrow(groups))) {

  this_condition <- groups$condition[i]
  this_sublib    <- groups$sublib[i]

  df_sub <- count_df_long %>%
    filter(condition == this_condition,
           sublib == this_sublib,
           count > 0) %>%                # needed for log2()
    mutate(log2_count = log2(count))

  p <- ggplot(df_sub, aes(x = log2_count)) +
    geom_histogram(bins = 50, fill = "blue", color = "white") +
    labs(
      title = paste("Condition:", this_condition, "| SubLib:", this_sublib),
      x = "log2(read count)",
      y = "Number of sgRNAs"
    ) +
    theme_bw() +
    theme(
      plot.margin = margin(t = 15, r = 10, b = 10, l = 10),
      plot.title = element_text(margin = margin(b = 10))
    ) +
    coord_cartesian(xlim = c(0, 15), ylim = c(0, 3000))
  print(p)  # important inside a loop
}

library(dplyr)
library(tibble)

coverage_df <- tibble(
  sample = c("I_L1_1", "I_L3_1", "I_L5_1",
             "L_L1_1", "L_L3_1", "L_L5_1",
             "U_L1_1", "U_L3_1", "U_L5_1"),
  coverage = c(100.00, 100.00, 100.00,
               99.99,  99.99,  99.96,
               99.99,  99.98,  99.99)
)
coverage_df <- tibble(
  sample = c("I_L1_1", "I_L3_1", "I_L5_1",
             "L_L1_1", "L_L3_1", "L_L5_1",
             "U_L1_1", "U_L3_1", "U_L5_1"),
  coverage = c(100.00, 100.00, 100.00,
               100.00,  100.00,  100.00,
               100.00,  100.00,  100.00)
)
coverage_df <- tibble(
  sample = c("I_L1_1", "I_L3_1", "I_L5_1",
             "L_L1_1", "L_L3_1", "L_L5_1",
             "U_L1_1", "U_L3_1", "U_L5_1"),
  coverage = c(100.00, 100.00, 100.00,
               100.00,  100.00,  99.99,
               99.99,  100.00,  100.00)
)
ggplot(coverage_df, aes(x = sample, y = coverage)) +
  geom_col(fill = "blue") +
  labs(
    title = "Correlation coverage",
    x = "Sample",
    y = "sgRNA coverage (%)"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(ylim = c(99.9, 100))   # zoom in so you can see differences

library(dplyr)
library(ggplot2)

groups <- count_df_long %>%
  distinct(condition, sublib)

for (i in seq_len(nrow(groups))) {

  this_condition <- groups$condition[i]
  this_sublib    <- groups$sublib[i]

  rank_df <- count_df_long %>%
    filter(condition == this_condition,
           sublib == this_sublib,
           count > 0) %>%                 # needed for log scale
    arrange(desc(count)) %>%              # highest -> lowest
    mutate(rank = row_number()) %>%       # rank index
    select(rank, count)

  p <- ggplot(rank_df, aes(x = rank, y = count)) +
    geom_line(color = "blue") +
    labs(
      title = paste("Knee plot of sgRNA read counts |", this_condition, "|", this_sublib),
      x = "Rank (sorted by count)",
      y = "Count (log10)"
    ) +
    scale_y_log10() +
    theme_bw() +
    theme(
      plot.margin = margin(t = 15, r = 10, b = 10, l = 10),
      plot.title = element_text(margin = margin(b = 10))
    )

  print(p)
}

```

``` {r Liangfu_Plots_1, eval=FALSE}
lg2fc_df <- maude_counts_df %>%
    left_join(merged_sgRNA_df %>% select(sgrna_id, Gene, count), by = c("sgRNA" = "sgrna_id")) %>% 
  rename(plasmid_count = count) %>% 
  filter(!isNontargeting)


vs_input <- plot_l2fc_scatter(lg2fc_df,
                 baseline = "input",
                 x_axis = "upper",
                 y_axis = "lower",
                 save_image_file_path = NULL,
                 genes = c("EGFP","AGO2"),
                 mark_N_special = 8,
                 pseudocount = 1e-9)
  
vs_plasmid <- plot_l2fc_scatter(lg2fc_df,
                 baseline = "plasmid_count",
                 x_axis = "upper",
                 y_axis = "lower",
                 save_image_file_path = NULL,
                 genes = c("EGFP","AGO2"),
                 mark_N_special = 8,
                 pseudocount = 1e-9)

csv_file_path <- get_file_path(results_output_folder,"L2FC_vs_input.csv")
write_csv(vs_input$means, csv_file_path)
csv_file_path <- get_file_path(results_output_folder,"L2FC_vs_plasmid.csv")
write_csv(vs_plasmid$means, csv_file_path)
```

```{r top_guide_per_gene, eval=FALSE}
dual_rep_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/"
all_conjugates_overlap_classified <- readRDS(file.path(dual_rep_dir, "all_conjugates_overlap_classified.rds"))
merged_sgRNA_df <- readRDS("/g/steinmetz/link/Amplicon_barcode_analysis/library/whole_genome_sublibs_1_to_4_merged_sgRNA_df.rds")
PA_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_bcwithqc"
DCA_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_bcwithqc"
GAL_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_bcwithqc"

guide_summary_all <- build_conjugate_maude_summary(
  conjugates_overlap = all_conjugates_overlap_classified,
  merged_sgRNA_df = merged_sgRNA_df,
  DCA_dir = DCA_dir,
  PA_dir = PA_dir,
  GAL_dir = GAL_dir,
  mode = "use_all"
)
best_sgRNA_all <- guide_summary_all$modified_overlap_df

guide_summary_confirmed <- build_conjugate_maude_summary(
  conjugates_overlap = all_conjugates_overlap_classified,
  merged_sgRNA_df = merged_sgRNA_df,
  DCA_dir = DCA_dir,
  PA_dir = PA_dir,
  GAL_dir = GAL_dir,
  mode = "use_only_confirmed_hits"
)
best_sgRNA_confirmed_hits <- guide_summary_confirmed$modified_overlap_df

saveRDS(best_sgRNA_all,"/g/steinmetz/link/sharing/nico/best_guides_all.rds")
saveRDS(best_sgRNA_confirmed_hits,"/g/steinmetz/link/sharing/nico/best_guides_confirmed_hits.rds")
```


```{r top_guide_mean_Z, eval=FALSE}
# dual_rep_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep/"
# all_conjugates_overlap_classified <- readRDS(file.path(dual_rep_dir, "all_conjugates_overlap_classified.rds"))
merged_sgRNA_df <- readRDS("/g/steinmetz/link/Amplicon_barcode_analysis/library/whole_genome_sublibs_1_to_4_merged_sgRNA_df.rds")
PA_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_bcwithqc"
DCA_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_bcwithqc"
GAL_dir <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_bcwithqc"

guide_z_ranks <- build_maude_guide_z_rank_df(
  merged_sgRNA_df = merged_sgRNA_df,
  DCA_dir = DCA_dir,
  PA_dir = PA_dir,
  GAL_dir = GAL_dir
)

guide_z_definitive_ranks <- guide_z_ranks %>%
  group_by(gene, entrez, sgRNA, seq) %>%
  summarise(
    mean_Z_across_screens = mean(mean_Z, na.rm = TRUE),
    n_screens = sum(!is.na(mean_Z)),
    screens_used = paste(sort(origin[!is.na(mean_Z)]), collapse = ";"),
    .groups = "drop"
  ) %>%
  mutate(
    mean_Z_across_screens = if_else(
      n_screens == 0,
      NA_real_,
      mean_Z_across_screens
    ),
    abs_mean_Z_across_screens = abs(mean_Z_across_screens),
    Z_direction_across_screens = case_when(
      mean_Z_across_screens > 0 ~ "positive",
      mean_Z_across_screens < 0 ~ "negative",
      mean_Z_across_screens == 0 ~ "zero",
      TRUE ~ NA_character_
    )
  ) %>%
  group_by(gene) %>%
  mutate(
    definitive_rank = dense_rank(desc(abs_mean_Z_across_screens))
  ) %>%
  ungroup() %>%
  arrange(gene, definitive_rank, sgRNA)

saveRDS(guide_z_ranks,"/g/steinmetz/link/sharing/nico/bcwithqc_guides_ranked_by_Z_score_full.rds")
writexl::write_xlsx(guide_z_ranks, "/g/steinmetz/link/sharing/nico/bcwithqc_guides_ranked_by_Z_score_full.xlsx")
saveRDS(guide_z_definitive_ranks,"/g/steinmetz/link/sharing/nico/bcwithqc_guides_ranked_by_Z_score.rds")
writexl::write_xlsx(guide_z_definitive_ranks, "/g/steinmetz/link/sharing/nico/bcwithqc_guides_ranked_by_Z_score.xlsx")

guide_z_ranks_old <- readRDS("/g/steinmetz/link/sharing/nico/best_guides/by_Z_score/align_UMI_tools_guides_ranked_by_Z_score_full.rds")
```


```{r cleanup, eval=FALSE}
delete_this = "/g/steinmetz/link/Amplicon_barcode_analysis/PA_subsampeling"
delete_bam_and_reads(root = delete_this, dry_run = FALSE)
```


```{r bcwithqc_time_compare, eval=FALSE}
res <- compare_log_runtimes(
  old_time = c(
    PA = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_bcwithqc_old/logs",
    # DCA = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_bcwithqc_old/logs",
    GAL = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_bcwithqc_old/logs"
  ),
  new_time = c(
    PA = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_bcwithqc/logs",
    # DCA = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_bcwithqc/logs",
    GAL = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_bcwithqc/logs"
  )
)

t_diff_fpairs <- res$file_pairs
t_diff_spairds <- res$sample_pairs
t_diff_summary <- res$summary


PA_mean_hits_bcwithqc <- automate_calc_replicate_means(
    file_info_suffix = "bcwithqc_reads_control_median_RI_comb_sample_comb_none",
    rds_output_folder = "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_bcwithqc/rds"
) 
```

```{r load_guide_stats, eval=FALSE}
pa_bc_gene_fpath <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_PA_bcwithqc/rds/MAUDE_gene_stats.rds"
check_pa_bc_gene_fpath <- readRDS(pa_bc_gene_fpath)
dca_bc_gene_fpath <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_DCA_bcwithqc/rds/MAUDE_gene_stats.rds"
check_dca_bc_gene_fpath <- readRDS(dca_bc_gene_fpath)
gal_bc_gene_fpath <- "/g/steinmetz/link/Amplicon_barcode_analysis/HepG2_dual_rep_GALNAC_bcwithqc/rds/MAUDE_gene_stats.rds"
check_gal_bc_gene_fpath <- readRDS(gal_bc_gene_fpath)
```

```{r load_guide_stats, eval=FALSE}
# R/R_functions/EA_00_run_plot.R

# 
# count_df_long <- readRDS(cfg$paths$count_df_fpath) %>%
#   rename(bin_name = condition)
# saveRDS(count_df_long, cfg$paths$count_df_fpath)
# read_count_df <- readRDS(cfg$paths$reads_count_df_fpath) %>%
#   rename(bin_name = condition)
# 
# saveRDS(read_count_df, cfg$paths$reads_count_df_fpath)
count_df_long <- readRDS(cfg$paths$count_df_fpath)
cfg$counting$data_type <- "umis"
if (identical("umis", cfg$counting$data_type)){
  run_umi_read_scatter_plots(
    umi_counts_df = count_df_long,
    read_counts_df = readRDS(cfg$paths$reads_count_df_fpath),
    cfg = cfg)
}
  
run_count_density_plots(count_df_long = count_df_long, cfg = cfg)

```

```{r Jia_analysis, eval=FALSE}
jia_read_counts <- readRDS("/g/steinmetz/link/Amplicon_barcode_analysis/Jia_CM_prelim/rds/count_df.rds")
jia_umi_counts <- readRDS("/g/steinmetz/link/Amplicon_barcode_analysis/Jia_CM_prelim/rds_umis/count_df.rds")
library <- readRDS(cfg$paths$library_path)
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(stringr)
library(patchwork)
library(tidytext)

# Helper: guess the sgRNA column in a table
guess_sg_col <- function(df) {
  candidates <- c("sgRNA", "sgrna_id", "sgrna_seq", "sgrna")
  hit <- intersect(candidates, colnames(df))
  if (length(hit) == 0) stop("No sgRNA column found in data frame.")
  hit[1]
}

# Helper: guess the count column in a table
guess_count_col <- function(df) {
  candidates <- c("count", "counts", "value")
  hit <- intersect(candidates, colnames(df))
  if (length(hit) == 0) stop("No count column found in data frame.")
  hit[1]
}

read_sg_col <- guess_sg_col(jia_read_counts)
umi_sg_col  <- guess_sg_col(jia_umi_counts)
read_ct_col  <- guess_count_col(jia_read_counts)
umi_ct_col   <- guess_count_col(jia_umi_counts)

# Make a common key for joining library <-> count tables.
# Prefer sgrna_id if available; otherwise fall back to sgRNA or sgrna_seq.
library_key_col <- guess_sg_col(library)

# Standardize columns
reads_df <- jia_read_counts %>%
  transmute(
    sgRNA = .data[[read_sg_col]],
    read_count = .data[[read_ct_col]],
    sample = if ("sample" %in% colnames(.)) sample else NA_character_,
    condition = if ("condition" %in% colnames(.)) condition else NA_character_
  )

umis_df <- jia_umi_counts %>%
  transmute(
    sgRNA = .data[[umi_sg_col]],
    umi_count = .data[[umi_ct_col]],
    sample = if ("sample" %in% colnames(.)) sample else NA_character_,
    condition = if ("condition" %in% colnames(.)) condition else NA_character_
  )

# If your library table uses sgrna_id, this join should work directly.
# If not, change `sgRNA` to the appropriate key column in library.
library_df <- library %>%
  rename(
    sgRNA = all_of(library_key_col)
  ) %>%
  select(sgRNA, gene, entrez_id, type, sgrna_seq)

# Join read + UMI counts and add gene
combined_df <- full_join(reads_df, umis_df, by = c("sgRNA", "sample", "condition")) %>%
  left_join(library_df, by = "sgRNA") %>%
  mutate(
    read_count = replace_na(read_count, 0),
    umi_count  = replace_na(umi_count, 0)
  )

# Optional: log transform for nicer plots
combined_df <- combined_df %>%
  mutate(
    log10_read = log10(read_count + 1),
    log10_umi  = log10(umi_count + 1)
  )

# ------------------------------------------------------------
# 1) Scatter plot: read count vs UMI count by sgRNA
# ------------------------------------------------------------
p_scatter <- ggplot(combined_df, aes(x = read_count, y = umi_count)) +
  geom_point(alpha = 0.35, size = 1) +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    title = "Read count vs UMI count by sgRNA",
    x = "Read count (log10 scale)",
    y = "UMI count (log10 scale)"
  ) +
  theme_bw()

# Optional: add correlation label
cor_val <- cor(combined_df$read_count, combined_df$umi_count, method = "spearman", use = "complete.obs")
p_scatter <- p_scatter +
  annotate("text", x = Inf, y = Inf, label = paste0("Spearman rho = ", round(cor_val, 3)),
           hjust = 1.1, vjust = 1.5, size = 4)

# ------------------------------------------------------------
# 2) Separate distributions of read and UMI counts
# ------------------------------------------------------------
dist_df <- combined_df %>%
  select(sgRNA, gene, read_count, umi_count) %>%
  pivot_longer(
    cols = c(read_count, umi_count),
    names_to = "count_type",
    values_to = "count"
  ) %>%
  mutate(
    count_type = recode(count_type,
                        read_count = "Read count",
                        umi_count = "UMI count")
  )

p_dist <- ggplot(dist_df, aes(x = count + 1, fill = count_type)) +
  geom_density(alpha = 0.35) +
  scale_x_log10() +
  labs(
    title = "Count distributions",
    x = "Count + 1 (log10 scale)",
    y = "Density",
    fill = ""
  ) +
  theme_bw()

# ------------------------------------------------------------
# 3) Lowest 10 sgRNAs by read / UMI count
# ------------------------------------------------------------

low20_sg_df <- combined_df %>%
  select(sgRNA, gene, read_count, umi_count) %>%
  distinct() %>%
  pivot_longer(
    cols = c(read_count, umi_count),
    names_to = "count_type",
    values_to = "count"
  ) %>%
  mutate(
    count_type = recode(count_type,
                        read_count = "Read count",
                        umi_count = "UMI count"),
    label = sgRNA
  ) %>%
  group_by(count_type) %>%
  slice_min(order_by = count, n = 10, with_ties = FALSE) %>%
  ungroup()

p_low20_sg <- ggplot(
  low20_sg_df,
  aes(x = tidytext::reorder_within(label, count, count_type), y = count, fill = count_type)
) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~count_type, scales = "free_y", ncol = 1) +
  tidytext::scale_x_reordered() +
  labs(
    title = "Lowest 10 sgRNAs by count type",
    x = NULL,
    y = "Count"
  ) +
  theme_bw()


# ------------------------------------------------------------
# 4) Lowest 10 genes by aggregated read / UMI count
# ------------------------------------------------------------

gene_df <- combined_df %>%
  filter(!is.na(gene), gene != "") %>%
  filter(type == "targeting") %>%
  group_by(gene) %>%
  summarise(
    read_count = sum(read_count, na.rm = TRUE),
    umi_count  = sum(umi_count, na.rm = TRUE),
    n_sgRNAs   = n_distinct(sgRNA),
    .groups = "drop"
  )

low20_gene_df <- gene_df %>%
  pivot_longer(
    cols = c(read_count, umi_count),
    names_to = "count_type",
    values_to = "count"
  ) %>%
  mutate(
    count_type = recode(count_type,
                        read_count = "Read count",
                        umi_count = "UMI count")
  ) %>%
  group_by(count_type) %>%
  slice_min(order_by = count, n = 10, with_ties = FALSE) %>%
  ungroup()

p_low20_gene <- ggplot(
  low20_gene_df,
  aes(x = tidytext::reorder_within(gene, count, count_type), y = count, fill = count_type)
) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~count_type, scales = "free_y", ncol = 1) +
  tidytext::scale_x_reordered() +
  labs(
    title = "Lowest 10 targeting genes by count type",
    x = NULL,
    y = "Aggregated count"
  ) +
  theme_bw()

# ------------------------------------------------------------
# Combine and print
# ------------------------------------------------------------

print(p_scatter)
print(p_dist)
print(p_low20_sg)
print(p_low20_gene)
```